---
title: "Profile-aware phase fuel-burn references for international benchmarking"
subtitle: "Exploratory evidence, study design, and validation gates for AICAP 2027"
author:
  - name: Lingling Ma
    affiliation: Civil Aviation University of China
  - name: Wei Dai
    affiliation: Civil Aviation University of China
  - name: Rainer Koelle
    affiliation: EUROCONTROL Performance Review Unit
date: today
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
  docx: default
  pdf: default
---

```{r}
#| label: setup
#| include: false
source(here::here("R", "setup.R"))
library(arrow)
library(dplyr)
library(ggplot2)
library(knitr)
library(tidyr)
```

```{r}
#| label: load-and-harmonize-data
#| cache: false

data_store <- Sys.getenv("FUELBURN_DATA_STORE", unset = here::here(".fuelburn-data-store"))

processed_path <- function(...) {
  file.path(data_store, ...)
}

eur_milestones <- arrow::read_parquet(
  processed_path("processed", "eur", "EUR-canonical-milestones-summer2025.parquet")
)
eur_segments_raw <- arrow::read_parquet(
  processed_path("processed", "eur", "EUR-level-segments-summer2025.parquet")
)
chn_milestones <- arrow::read_parquet(
  processed_path("processed", "chn", "CHN-canonical-milestones-summer2025.parquet")
)
chn_segments_raw <- arrow::read_parquet(
  processed_path("processed", "chn", "CHN-level-segments-summer2025.parquet")
)
chn_phase_summaries <- arrow::read_parquet(
  processed_path("processed", "chn", "CHN-phase-summaries-summer2025.parquet")
)

altitude_band_from_ft <- function(x) {
  case_when(
    is.na(x) ~ NA_character_,
    x < 10000 ~ "below_FL100",
    x < 18000 ~ "FL100_FL180",
    x < 24000 ~ "FL180_FL240",
    TRUE ~ "above_FL240"
  )
}

standardise_level_segments <- function(data, region) {
  data |>
    mutate(
      REGION = region,
      HAS_END = if ("HAS_END" %in% names(data)) .data$HAS_END else !is.na(.data$END_TIME),
      FUEL_BURN_KG = if ("FUEL_BURN_KG" %in% names(data)) .data$FUEL_BURN_KG else .data$FUEL_KG,
      MEAN_ALTITUDE_FT = if ("MEAN_ALTITUDE_FT" %in% names(data)) {
        .data$MEAN_ALTITUDE_FT
      } else {
        (.data$START_ALT_FT + .data$END_ALT_FT) / 2
      },
      ALTITUDE_BAND = if ("ALTITUDE_BAND" %in% names(data)) {
        .data$ALTITUDE_BAND
      } else {
        altitude_band_from_ft(.data$MEAN_ALTITUDE_FT)
      },
      LEVEL_CONTEXT_PHASE = if ("LEVEL_CONTEXT_PHASE" %in% names(data)) {
        .data$LEVEL_CONTEXT_PHASE
      } else {
        .data$PHASE_CONTEXT
      },
      qc_flag = if ("qc_flag" %in% names(data)) .data$qc_flag else "not_assessed",
      duration_min = .data$DURATION_SEC / 60,
      kg_min = .data$FUEL_BURN_KG / .data$duration_min,
      altitude_band = factor(
        .data$ALTITUDE_BAND,
        levels = c("below_FL100", "FL100_FL180", "FL180_FL240", "above_FL240")
      )
    )
}

eur_segments <- standardise_level_segments(eur_segments_raw, "EUR")
chn_segments <- standardise_level_segments(chn_segments_raw, "CHN")
level_segments <- bind_rows(eur_segments, chn_segments)

milestones <- bind_rows(
  eur_milestones |> mutate(REGION = "EUR"),
  chn_milestones |> mutate(REGION = "CHN")
)

dataset_summary <- tibble::tibble(
  REGION = c("EUR", "CHN"),
  flights = c(n_distinct(eur_milestones$SOURCE_UID), n_distinct(chn_milestones$SOURCE_UID)),
  milestones = c(nrow(eur_milestones), nrow(chn_milestones)),
  level_segments = c(nrow(eur_segments), nrow(chn_segments)),
  source_qc_ok_segments = c(sum(eur_segments$qc_flag == "ok", na.rm = TRUE), sum(chn_segments$qc_flag == "ok", na.rm = TRUE))
) |>
  mutate(source_qc_ok_share = source_qc_ok_segments / level_segments)

eur_flights <- dataset_summary$flights[dataset_summary$REGION == "EUR"]
chn_flights <- dataset_summary$flights[dataset_summary$REGION == "CHN"]
eur_n_segments <- dataset_summary$level_segments[dataset_summary$REGION == "EUR"]
chn_n_segments <- dataset_summary$level_segments[dataset_summary$REGION == "CHN"]
```

# Executive Summary

The study aims to establish an enriched, phase-dependent fuel-burn lookup table for international operational performance benchmarking. It connects operator-recorded Chinese QAR fuel flow with stakeholder-endorsed European AEM estimates applied to Network Manager trajectories through a common analytical contract. Harmonisation concerns the meaning of events, intervals, units, and eligibility; identical raw processing is neither available nor required.

The intended contribution is to test whether operational profile descriptors improve fuel-reference estimates beyond one value per aircraft type and phase. Climb and descent smoothness, cruise altitude exposure, and TOC-to-TOD duration are related but distinct questions. A smooth climb or descent is a candidate low-interruption profile; sustained level cruise is expected. A low fuel-flow percentile is not automatically an efficient profile.

**Evidence status, 9 October 2026:** the local release contains `r formatC(eur_flights, format="d", big.mark=",")` EUR flights and `r formatC(chn_flights, format="d", big.mark=",")` CHN flights. File hashes match the manifest, but new interval and distance checks identify unresolved analytical defects. The descriptive figures below are diagnostics of delivered records. No validated international coefficients, regional efficiency ranking, or achievable fuel savings are claimed.

**Working documents:** this exploratory paper develops the argument and initial evidence; `technical-note-data-preparation.qmd` describes source preparation; `technical-note-methodology.qmd` records analytical decisions and validation studies; `notes/study-worker-plan.md` provides work packages and acceptance gates.

# 1. Foundation and research contribution

## 1.1 What the 2025 paper established

The supplied ICNS2025 paper, *Fuel flow analysis for different flight phases based on a data-driven approach* (Ma, Dai, and Koelle), established a useful milestone architecture and an initial CHN/EUR comparison. Its reported medium-aircraft full-flight regression slopes were 35.55 kg/min in China and 34.45 kg/min in Europe. These are sample-specific regression results, not phase-specific lookup coefficients or validation targets. Its preliminary airborne benefit-pool estimate below 200 kg was explicitly qualified and depended on a 20th-percentile reference and aggregation choices.

The previous study used Chinese October 2023 data and European summer 2024 data. The present summer 2025 release offers substantially more Chinese observations and a closer seasonal comparison, subject to actual date and coverage checks. The previous paper also identified unresolved profile variability, heuristic TOC/TOD detection, missing weight/load information, and problematic heavy-aircraft aggregation. Those limitations motivate validation rather than assuming that earlier numerical results should recur.

The legacy partition is TXO (AOBT–ATOT), CLIMB100 (ATOT–D100), ENR100 (D100–A100), DESC100 (A100–ALDT), and TXI (ALDT–AIBT). ENR100 is a distance-defined interval and can contain climbing and descending flight. It must remain distinct from TOC–TOD cruise/enroute analysis.

## 1.2 Questions and testable contributions

| Question | Estimand or comparison | Evidence needed |
|---|---|---|
| RQ1: Does climb/descent interruption describe fuel variation? | Conditional phase fuel and fuel rates versus level time, count, and altitude exposure | Validated phase windows, complete profile coverage, and common detector qualification |
| RQ2: Does enrichment improve a practical lookup? | Held-out error of type × phase baseline versus profile-aware cells | Adequate independent flight support; all tuning confined to development data |
| RQ3: How do cruise level and TOC–TOD duration relate to fuel? | Within-type/route associations, exposure-specific rates, and whole-airborne fuel | Validated distance, TOC/TOD, altitude exposure, and available confounders |
| RQ4: Can coefficients transfer between sources? | Source-specific estimates on common support; transfer error | Matched context and explicit measurement/model provenance |

RQ1 and RQ2 form the primary conference contribution. RQ3 is a focused side study that can enter the paper if validated; RQ4 establishes the limits of international transfer. The design can yield a useful negative finding if enrichment fails to improve prediction or common support is insufficient.

# 2. Analytical framework

## 2.1 Source-specific preparation, common analytical contract

China integrates recorded engine fuel flow; Europe uses AEM model estimates on NM trajectories. Source and region are therefore confounded in this sample. A CHN–EUR coefficient difference cannot identify a regional operational effect separately from measurement/model bias, fleet, route, and operator selection. Where feasible, applying a comparable model configuration to a QAR trajectory provides a bridge study on the same flight. Agreement with the EUR model's own output is an implementation check, not independent validation.

The analysis contract consists of:

1. **Flights:** one row per `(REGION, SOURCE_UID)`; verified type, route direction, dates, source configuration, coverage, and eligibility.
2. **Milestones:** one row per event role; timestamp, cumulative consumed fuel, cumulative flown distance, altitude reference, and derivation provenance. Several different roles may share a physical trajectory point.
3. **Phase intervals:** one row per flight and explicitly named phase scheme; ordered boundaries and positive exposure, with fuel and distance differences.
4. **Level intervals and profile descriptors:** unique paired intervals, phase clipping, non-overlapping time exposure, and independent fuel/distance eligibility flags.

The same column names alone do not meet this contract. The present CHN and EUR schemas differ, and the production EUR milestone table still contains legacy `LVL` and generic `FL100` events.

## 2.2 Phase and milestone semantics

The physical airborne partition is ATOT–TOC (climb), TOC–TOD (enroute), and TOD–ALDT (descent). Retain the 2025 distance partition as a separate reproducibility sensitivity. Never add intervals from these two partitions together. Zero-duration cruise is an explicit short-flight/no-cruise case, not an ordinary rate observation. Missing or unordered boundaries are flagged; missing landing milestones are not silently replaced by in-block times.

Along-track D040/D100/D200 and A200/A100/A040 anchors support a distance partition only after cumulative distance is validated. Geometric distances to airports are separate fields and may be used for a 200 NM airport window. Directional FL crossings are analytical pressure-altitude crossings. AGL windows require field elevation; QNH and pressure altitude must be distinguished, particularly for high-elevation airports such as ZPPP. Rich anchors are diagnostic options, not all mandatory lookup dimensions.

TOC/TOD rules must distinguish capped initial cruise, later step climbs, and final sustained descent. Their provenance and alternative definitions must be retained. Common phase context should be based on the agreed milestone windows; altitude-only labelling cannot decide that a level-off above FL200 belongs to cruise.

## 2.3 Smoothness and profile classes

Level detection and reporting qualification are separate steps. Published EUROCONTROL CCO/CDO work provides a 300 ft/min reference lineage; it does not validate every local duration, smoothing, or altitude-band choice. Reprocess a representative QAR sample at European observation cadence to quantify how sampling affects short level-segment detection.

For flight i and phase p, measure qualified level time as the length of the **union** of intervals clipped to that phase:

$$L_{ip}=\left|\bigcup_j ([s_{ij},e_{ij})\cap[a_{ip},b_{ip}))\right|,\qquad S_{ip}=L_{ip}/T_{ip}.$$

Record count, longest interruption, absolute level time, level share, and altitude residence. Use only validated detection coverage to distinguish a true zero from missing evidence. Union time is an audit safeguard; it does not repair incorrect interval pairing or fuel totals. Clip fuel from cumulative fuel at boundaries, with a declared interpolation rule; do not allocate a long interval's fuel by time without checking that approximation.

For climb and descent, initially call classes **low, typical, and high interruption**. Operational thresholds or development-sample quantiles may define them, but never use the target fuel outcome to assign classes. A no-qualified-level category can be retained where sample support permits. Rename a class “highly efficient” only if matched, out-of-sample fuel evidence supports that interpretation. For cruise, describe altitude exposure and step changes: a long level segment is normal and must not trigger an inefficient label.

## 2.4 What the lookup estimates

A phase's total fuel is not its fuel rate. For valid fuel F, time T, and flown distance D:

$$r_t=F/T\quad[\mathrm{kg/min}],\qquad r_d=F/D\quad[\mathrm{kg/NM}].$$

A typical-flight median rate and the exposure-weighted coefficient $\sum F/\sum T$ answer different questions. Publish the aggregation rule. Count independent flights, not only segments; flights with many fragments should not dominate a lookup. Report total phase fuel separately to assess whether low-interruption profiles also have lower matched consumption.

The initial lookup key is **source/region × exact aircraft type × phase × profile class**, with stage length as a continuous adjustment or a small number of validated bands. Add altitude only where it improves held-out accuracy and support remains sufficient. Aircraft-family fallback requires evidence; A320 and A20N are separate primary types. Taxi requires a separate ground-operation design and is outside the first airborne release.

A coefficient multiplied by extra minutes estimates a scenario conversion under the cell's operating conditions. It is not the marginal causal cost of a level-off: alternative flight paths change thrust, speed, altitude, distance, and the distribution of time across phases.

## 2.5 Percentiles are candidates, not efficiency labels

P20/P50/P80 describe the conditional distribution inside a defined cell. They are neither confidence limits nor best/standard/worst operational classes. P50–P20 is a distributional spread, not proven achievable savings. P20 fuel rate can select an aircraft descending at low thrust; P20 phase fuel can select a shorter interval, lower weight, or favourable winds. Choosing P20 does not eliminate wind effects.

Compare median and exposure-weighted coefficients with conditional percentiles at P10/P20/P25 and P75/P80/P90. Assess stability, coverage, and held-out prediction or quantile calibration. A bootstrap confidence interval on a percentile estimates sampling uncertainty; the percentile range describes outcome variability. Neither includes unidentified model bias. A percentile-based reference remains optional if it adds no defensible operational meaning.

## 2.6 Cruise altitude, duration, and the spurious-results study

The hypothesis that lower cruise levels may consume more fuel is operationally plausible, but its sign and magnitude are conditional on aircraft type, mass, speed, winds, stage length, and attainable levels. Test the association rather than imposing a monotonic lookup. Higher is not universally better.

For TOC–TOD retain total fuel, time, flown distance, kg/min, kg/NM, altitude residence, time-weighted altitude, and step changes. Report altitude summaries from validated points or true source intervals; `(TOC altitude + TOD altitude)/2` is not a trajectory mean. Typical maximum altitude is a diagnostic, not a surrogate for exposure.

Separate three comparisons: fixed exposure fuel-rate associations; observed total cruise fuel with duration and distance; and whole-airborne fuel ATOT–ALDT. A longer cruise phase can mean a longer route, but also earlier TOC or later TOD replacing high-thrust climb or powered descent. With $F_{air}=F_{climb}+F_{enroute}+F_{descent}$, an isolated cruise reduction need not reduce whole-flight fuel. Time, distance, and speed are related; avoid treating all as independent controls in an unrestricted regression.

Analyse within type and route direction first, with date/batch and available mass, speed, and weather information. Publish pooled and within-group patterns to reveal mixture reversals. An altitude/duration response surface is a diagnostic if there is enough support; do not extrapolate to levels or durations absent from a route/type cell. Without requested/optimal altitude or a credible counterfactual, observed level differences are not a quantified ATM-imposed altitude penalty.

# 3. EUR and CHN Data Exploration

## 3.1 Dataset Characteristics

**Processed Data Summary (Summer 2025):**

- **EUR:** `r formatC(eur_flights, format="d", big.mark=",")` flights, `r formatC(nrow(eur_milestones), format="d", big.mark=",")` canonical milestone rows, `r formatC(eur_n_segments, format="d", big.mark=",")` level segments.
- **CHN:** `r formatC(chn_flights, format="d", big.mark=",")` flights, `r formatC(nrow(chn_milestones), format="d", big.mark=",")` canonical milestone rows, `r formatC(chn_n_segments, format="d", big.mark=",")` level segments.
- **Data source:** R2 manifest production products under `processed/eur` and `processed/chn`.
- **Purpose:** establish common exploratory descriptors before building fuel-burn lookup tables.

```{r}
#| label: dataset-summary
#| tbl-cap: "Release inventory; source QC labels do not establish analytical validity"

dataset_summary |>
  mutate(source_qc_ok_share = 100 * source_qc_ok_share) |>
  kable(
    digits = 1,
    col.names = c("Region", "Flights", "Milestone rows", "Level segments", "Source QC ok rows", "Source QC ok share [%]"),
    format.args = list(big.mark = ",")
  )
```

**Milestone counts by region:**

```{r}
#| label: milestone-counts
#| tbl-cap: "Top canonical milestone labels by region"

milestones |>
  count(REGION, MST, sort = TRUE) |>
  group_by(REGION) |>
  slice_max(n, n = 12, with_ties = FALSE) |>
  ungroup() |>
  arrange(REGION, desc(n)) |>
  kable(
    col.names = c("Region", "Milestone", "Count"),
    format.args = list(big.mark = ",")
  )
```

## 3.2 Level-Segment Duration Distributions

```{r}
#| label: duration-by-phase
#| fig-cap: "Level-segment duration by region and phase context"
#| fig-width: 10
#| fig-height: 6
#| warning: false

level_segments |>
  filter(HAS_END, DURATION_SEC > 0, !is.na(LEVEL_CONTEXT_PHASE)) |>
  filter(LEVEL_CONTEXT_PHASE %in% c("CLIMB", "ENROUTE", "DESCENT")) |>
  ggplot(aes(x = duration_min, fill = REGION)) +
  geom_histogram(bins = 50, alpha = 0.7, position = "identity") +
  facet_grid(REGION ~ LEVEL_CONTEXT_PHASE, scales = "free_y") +
  scale_x_continuous(limits = c(0, 60), breaks = seq(0, 60, 10)) +
  labs(
    x = "Duration [minutes]",
    y = "Number of segments",
    fill = "Region"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

**Observations:**

```{r}
#| label: duration-summary

duration_summary <- eur_segments |>
  filter(HAS_END, DURATION_SEC > 0, !is.na(LEVEL_CONTEXT_PHASE)) |>
  mutate(REGION = "EUR") |>
  bind_rows(
    chn_segments |>
      filter(HAS_END, DURATION_SEC > 0, !is.na(LEVEL_CONTEXT_PHASE)) |>
      mutate(REGION = "CHN")
  ) |>
  filter(LEVEL_CONTEXT_PHASE %in% c("CLIMB", "ENROUTE", "DESCENT")) |>
  group_by(REGION, LEVEL_CONTEXT_PHASE) |>
  summarise(
    n = n(),
    mean = mean(duration_min),
    median = median(duration_min),
    p20 = quantile(duration_min, 0.2),
    p80 = quantile(duration_min, 0.8),
    .groups = "drop"
  )

kable(duration_summary, digits = 1,
      col.names = c("Region", "Phase context", "N", "Mean [min]", "Median [min]", "P20 [min]", "P80 [min]"),
      format.args = list(big.mark = ","))
```

These are distributions of the delivered interval rows. EUR repetitions and overlaps, unequal detection methods, and heuristic phase labels prevent an operational regional interpretation. The histogram displays durations up to 60 minutes; longer rows remain in the summary table. Recompute these plots from validated intervals before selecting results for the conference paper.

## 3.3 Altitude Band Analysis

```{r}
#| label: altitude-bands
#| fig-cap: "Level-segment distribution by altitude band"
#| fig-width: 10
#| fig-height: 5
#| warning: false

level_segments |>
  filter(HAS_END, DURATION_SEC > 0, !is.na(altitude_band)) |>
  count(REGION, altitude_band) |>
  ggplot(aes(x = altitude_band, y = n, fill = REGION)) +
  geom_col(position = "dodge") +
  geom_text(aes(label = scales::comma(n)), position = position_dodge(width = 0.9), vjust = -0.4, size = 3) +
  labs(
    x = "Altitude band",
    y = "Number of segments",
    fill = "Region"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

**Key findings:**

```{r}
#| label: altitude-summary

alt_summary <- level_segments |>
  filter(HAS_END, DURATION_SEC > 0, !is.na(altitude_band)) |>
  group_by(REGION, altitude_band) |>
  summarise(
    n = n(),
    median_dur = median(duration_min),
    .groups = "drop"
  ) |>
  group_by(REGION) |>
  mutate(pct = 100 * n / sum(n)) |>
  ungroup() |>
  select(REGION, altitude_band, n, pct, median_dur) |>
  arrange(REGION, altitude_band)

kable(alt_summary, digits = 1,
      col.names = c("Region", "Altitude band", "N", "% within region", "Median duration [min]"),
      format.args = list(big.mark = ","))
```

These bands are provisional reporting bins. Segment counts do not measure time exposure or establish optimal altitude. An endpoint-average altitude can conceal excursions within an interval; point-based residence times are needed before using altitude as a fuel-rate lookup dimension.

## 3.4 Aircraft Type Patterns

```{r}
#| label: type-analysis
#| fig-cap: "Level-segment duration by aircraft type"
#| fig-width: 10
#| fig-height: 6
#| warning: false

level_segments |>
  filter(HAS_END, DURATION_SEC > 0, !is.na(TYPE)) |>
  mutate(
    aircraft_group = case_when(
      str_detect(TYPE, "A319|A320|A321|A20N|A21N|B737|B738|B739|C919") ~ "Medium",
      TRUE ~ "Heavy/Other"
    )
  ) |>
  group_by(REGION, TYPE) |>
  filter(n() >= 250) |>
  ungroup() |>
  ggplot(aes(x = reorder(TYPE, duration_min, median), y = duration_min, fill = REGION)) +
  geom_boxplot() +
  coord_cartesian(ylim = c(0, 40)) +
  facet_wrap(~REGION, scales = "free_x") +
  labs(
    x = "Aircraft type",
    y = "Duration [minutes]",
    fill = "Region"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom", axis.text.x = element_text(angle = 45, hjust = 1))
```

**Aircraft-specific observations:**

```{r}
#| label: type-summary

type_summary <- level_segments |>
  filter(HAS_END, DURATION_SEC > 0, !is.na(TYPE)) |>
  group_by(REGION, TYPE) |>
  summarise(
    n = n(),
    median_dur = median(duration_min),
    p20_dur = quantile(duration_min, 0.2),
    p80_dur = quantile(duration_min, 0.8),
    .groups = "drop"
  ) |>
  group_by(REGION) |>
  slice_max(n, n = 8, with_ties = FALSE) |>
  ungroup() |>
  arrange(REGION, desc(n))

kable(type_summary, digits = 1,
      col.names = c("Region", "Type", "N", "Median [min]", "P20 [min]", "P80 [min]"),
      format.args = list(big.mark = ","))
```

- The common type set is not identical across the two samples, so type-specific lookup tables need a mapping layer for aircraft family/class.
- The Chinese sample includes additional domestic and long-haul types; the European sample is dominated by short/medium-haul narrowbodies.
- Type-specific stratification remains necessary before fuel-rate comparisons.

## 3.5 Quality Control Insights

```{r}
#| label: qc-flags
#| tbl-cap: "Level-segment QC flag distribution"

level_segments |>
  count(REGION, qc_flag, sort = TRUE) |>
  group_by(REGION) |>
  mutate(pct = 100 * n / sum(n)) |>
  ungroup() |>
  arrange(REGION, desc(n)) |>
  kable(digits = 1,
        col.names = c("Region", "QC Flag", "Count", "% within region"),
        format.args = list(big.mark = ","))
```

**QC implications:** `not_assessed` records the absence of delivered CHN QC flags. EUR `ok` is a source flag and does not test uniqueness or overlap. Neither label is a common eligibility rule. Endpoint altitude change also cannot establish within-interval stability. The readiness audit below takes precedence over these source flags.

# 4. Analytical readiness audit

The following audit reads the local manifest and production artifacts without changing them. It regenerates aggregate diagnostics under `outputs/study-design/`; these outputs contain no flight identifiers. Reproduce independently with `Rscript scripts/12-audit-study-readiness.R`.

```{r}
#| label: study-readiness-audit
#| include: false
invisible(capture.output(source(here::here("scripts", "12-audit-study-readiness.R"), local = TRUE)))
```

```{r}
#| label: study-readiness-table
#| tbl-cap: "Readiness checks on the hash-verified local release; interval rows are not independent flights"
study_readiness |>
  select(REGION, flights, segment_rows, distinct_segment_keys,
         repeated_segment_keys, flights_with_overlapping_intervals,
         intervals_implied_speed_over_700kt, flights_missing_unique_ALDT) |>
  pivot_longer(-REGION, names_to = "Check", values_to = "Value") |>
  pivot_wider(names_from = REGION, values_from = Value) |>
  kable(format.args = list(big.mark = ","))
```

All seven local artifacts listed in the manifest match their SHA-256 hashes. This verifies release identity, not scientific validity. Four findings require upstream resolution:

- **EUR interval identity and overlap:** 86,142 rows represent 47,849 distinct flight/segment keys; 29,631 keys repeat, adding 38,293 rows. Overlap occurs on 18,860 flights. Summed interval exposure is about 20,132 hours versus a union of about 12,802 hours, an inflation of approximately 57% relative to the union. These union hours are diagnostics, not corrected level time. Even `ok` rows can repeat. Inspection of `reconstruct_level_segments()` and `derive_level_segments()` suggests that implicit ends and joins on cumulative start order can yield several ends for one start. This is a candidate mechanism to trace against source segments, not a validated repair.
- **CHN distance integrity:** 17,955 of 64,981 intervals imply speeds above 700 kt from reported NM and seconds; the median TOC–TOD implied speed is approximately 830 kt. The 700 kt screen is an analyst diagnostic, not an operationally validated exclusion cutoff. Investigate coordinate units, jumps, cumulative-distance construction, time alignment, and batch provenance. Do not repair by assuming a universal unit multiplier. Distance-based matching, kg/NM, and distance-derived milestone placement depend on this check.
- **Missing boundaries:** CHN has 1,190 flights without ALDT, one without ATOT, and 22 without a unique TOC/TOD pair. Strict ATOT < TOC < TOD < ALDT holds for 5,220 CHN flights and 21,321 EUR flights. These counts are chronology diagnostics before other eligibility checks, not final study samples. Recover events from the source only under a justified shared definition; retain domain-specific samples and selection comparisons.
- **Semantics and actual schemas:** EUR milestones/segments have 21/31 columns; CHN has 27/22. EUR segments have no delivered segment-distance column. The EUR production milestone file has no `LVL_START`/`LVL_END` and retains generic `FL100`. The enrichment script matches uppercase phase names, while delivered start phases are `Lvl_climb`/`Lvl_descent`, then assigns many high-altitude intervals to ENROUTE. Both source phase assignment and canonical export lineage need review.

CHN AOBT dates cover June, July, and August 2025 (1,954 / 2,228 / 2,239 flights). Existing handover prose that leaves August uncertain is superseded by this field-level observation. Confirm date/time semantics with the source owner: EUR carries `Europe/Paris`, CHN `UTC`; different display zones alone do not establish a timestamp error.

## 4.1 Common support before matching

```{r}
#| label: aircraft-common-support
#| tbl-cap: "Shared exact types before quality filtering or distance/altitude matching"
type_support |>
  filter(EUR > 0, CHN > 0) |>
  slice_head(n = 12) |>
  kable(format.args = list(big.mark = ","))
```

B738, A320, A321, A20N, and A21N are plausible starting types, subject to quality and context overlap. Widebodies are a separate exploratory stratum because European support is much smaller. B737-MAX remains unresolved without its variant. C919 is CHN-only and cannot support direct international type comparisons. Same-region routes can support adjustment and validation; geographically different route labels cannot be paired as identical routes across regions.

Some earlier notes misname the Chinese airports: ZSSS is Shanghai Hongqiao, and ZLXY is Xi'an Xianyang. Use verified ICAO metadata in analysis. The audit reports ICAO codes to avoid propagation of those labels.

# 5. Lookup demonstration and validation design

## 5.1 Delivered-row diagnostic

The following table retains an initial exploration of the delivered data for traceability. It pools several aircraft types under a provisional medium label and uses the delivered ENROUTE context. EUR repeated intervals and CHN distance/boundary defects remain. These figures must not be distributed as lookup coefficients.

```{r}
#| label: example-lookup
#| tbl-cap: "Diagnostic summaries of delivered enroute-labelled rows; not validated lookup coefficients"
#| warning: false

level_segments |>
  filter(HAS_END, DURATION_SEC > 0, !is.na(TYPE), !is.na(altitude_band)) |>
  mutate(
    aircraft_group = case_when(
      str_detect(TYPE, "A319|A320|A321|A20N|A21N|B737|B738|B739|C919") ~ "Medium",
      TRUE ~ "Other"
    )
  ) |>
  filter(aircraft_group == "Medium", LEVEL_CONTEXT_PHASE == "ENROUTE") |>
  group_by(REGION, altitude_band) |>
  summarise(
    n = n(),
    dur_p20 = quantile(duration_min, 0.2),
    dur_p50 = quantile(duration_min, 0.5),
    dur_p80 = quantile(duration_min, 0.8),
    kg_min_p50 = median(kg_min, na.rm = TRUE),
    .groups = "drop"
  ) |>
  arrange(REGION, altitude_band) |>
  kable(digits = 1,
        col.names = c("Region", "Altitude band", "N", "Duration P20 [min]", "P50 [min]", "P80 [min]", "Median kg/min"),
        format.args = list(big.mark = ","))
```

## 5.2 Release schema and sparse-cell handling

A lookup release must include source, exact type, phase scheme, profile definition and version, stage-length support, any altitude dimension, metric and units, estimator, coefficient, percentile summaries where justified, independent flight count, route/date coverage, uncertainty, applicability range, and eligibility/fallback status. Keep source-specific estimates side by side until transfer is supported. A pooled coefficient needs an explicit weighting scheme and sensitivity to that choice.

Use a documented hierarchy: validated profile-specific cell; then the same type/phase cell without the optional profile dimension; then a tested family fallback; otherwise “insufficient support”. Never silently replace an unsupported cell with zero. Cell support thresholds and acceptable uncertainty are study design decisions to trial, not established constants.

## 5.3 Compare nested estimates out of sample

- **M0:** one coefficient per exact type and phase, with declared aggregation.
- **M1:** M0 plus validated stage-length context.
- **M2:** M1 plus climb/descent interruption class or cruise altitude exposure.

Use the same eligible held-out flights for all model comparisons. Separate development and evaluation by flight, date, and preferably route blocks; keep every segment of one flight in the same split. Derive bin edges, profile quantiles, and any tuning solely on development data. Report MAE, signed bias, and error by type/phase/source; avoid percentage errors for very small fuel totals. For conditional quantiles assess empirical held-out coverage. Cluster uncertainty by flight and examine route/day sensitivity; acknowledge that three Chinese route pairs limit population inference.

The decision is whether enrichment adds useful accuracy and interpretability without excessive sparsity. Accept M0/M1 if M2 does not pass. A transfer experiment trains in one source and evaluates in the other on common support, and reports the resulting bias rather than declaring either source ground truth.

# 6. Parameter trials and evidence boundaries

The current FL100/FL180/FL240 bins, distance bands, duration classes, and P20/P80 summaries are exploratory choices. Distinguish four statuses in the decision register: externally referenced definition; operational hypothesis; empirical candidate; and adopted after validation. A published threshold is applicable only with its original vertical reference, measurement window, and detection conditions.

Trial detector/reporting parameters on a small labelled sample first: vertical rate 200/300/400 ft/min; qualifying duration 20/30/60/120 seconds; within-interval altitude range 100/200/300 ft; and source-appropriate gap/smoothing settings. These values are proposed sensitivity scenarios, not endorsed settings. A 31-record smoother is not a fixed duration across sampling rates. A five-minute exclusion would remove many climb/descent interruptions and must not become the default without evidence.

For altitude, compare a continuous relationship with the current operational bands and narrower cruise bands only where data support permits. Shift boundaries by ±1,000/2,000 ft as sensitivity trials. For stage length, compare continuous adjustment and simple prespecified bands using validated flown distance, with geometric airport-pair distance as a separately named check. Do not optimise all dimensions or cut points on the full dataset. Store each trial's coverage, stability, and predictive impact in the Technical Note and freeze the selected configuration before final evaluation.

There are three distinct uncertainty sources: sampling error; sensitivity to definitions/parameters; and source-model/measurement uncertainty. A bootstrap handles the first, not the other two. Smoothness describes operational profiles but does not attribute them to ATM: terrain, procedures, speed management, weather, weight, and operator choices can all matter.

# 7. Conference paper and supporting Technical Note

The [organiser's EasyChair call](https://easychair.org/cfp/AICAP-2027), checked on 9 October 2026, specifies **eight pages including references, appendices, and other supplementary content**, IEEE double-column PDF, and a 6 MB file limit. The [official conference page](https://event.ntu.edu.sg/AICAP-2027) confirms submission on **30 October 2026**, acceptance notification on 1 December, and camera-ready submission on 15 December. The supplied NTU call-for-papers subpage was inaccessible during this review; the organiser's EasyChair listing supplies the detailed format requirements.

Treat this document as an expandable working paper. The conference manuscript should contain a concise motivation, the common analytical contract, the independent profile definition, one validated lookup demonstration, a nested-model comparison, the most consequential sensitivity result, and measured limitations. Reserve implementation detail, all parameter trials, bridge studies, missing-data diagnostics, and the cruise side study for the Technical Note unless they directly establish the main contribution. The external note must not be assumed to bypass submission rules or replace self-contained methods in the eight pages.

Indicative page allocation: title/abstract/introduction 1.0; data and source comparability 1.0; methods 1.5; principal results 2.0; validation/sensitivity 1.0; discussion/conclusion 0.75; references 0.75. This totals eight pages and must be checked in the actual IEEE template. The present Quarto PDF default is not the conference layout.

# 8. Conclusions and next decision

The defensible contribution is a test of phase fuel references enriched by independently defined operational context, with transparent cross-source limitations. The present release is reproducibly identifiable but needs interval, distance, milestone, and phase-semantic validation before coefficient estimation. P20/P50/P80 remain optional distribution summaries. Low-interruption profiles and lower cruise levels are hypotheses whose fuel implications must be demonstrated, rather than encoded in class labels.

The worker plan begins with release validation and source-owner checks, then builds phase/profile tables, trials parameter choices, compares lookup accuracy, and completes the altitude/duration side study. No numerical fuel-saving conclusion follows from the delivered-row plots in this exploratory draft.

# References and evidence

1. Ma, L., Dai, W., and Koelle, R. (2025). *Fuel flow analysis for different flight phases based on a data-driven approach*. Supplied ICNS2025 paper; numerical findings above are attributed to that document.
2. EUROCONTROL. [Continuous climb and descent operations](https://www.eurocontrol.int/concept/continuous-climb-and-descent-operations). Harmonised definitions and stakeholder context.
3. EUROCONTROL (2020). [European CCO/CDO Action Plan, Appendix E](https://www.eurocontrol.int/sites/default/files/2020-11/european-cco-cdo-action-plan-appendix-e.pdf). Examples of level-flight definitions and reporting windows; local qualification settings still require validation.
4. EUROCONTROL. [Advanced Emission Model](https://avenir-catalogue.eurocontrol.int/aem-details.html). Source-model context; record the configuration actually used in this study.
5. EUROCONTROL (2021). [Data Snapshot 21: Aircraft flying higher, more efficiently and sustainably](https://www.eurocontrol.int/publication/eurocontrol-data-snapshot-21-aircraft-flying-higher-more-efficiently-and-sustainably). Operational motivation for the altitude hypothesis, not proof of a universal altitude effect in this sample.
6. AICAP 2027. [Organiser's submission guidelines](https://easychair.org/cfp/AICAP-2027) and [official conference dates](https://event.ntu.edu.sg/AICAP-2027), accessed 9 October 2026.

**Status:** study design and delivered-release diagnostics, 9 October 2026. All parameter grids and work assignments are proposals pending study evidence and collaborator review.
