---
title: "Refined Fuel Burn Estimation Framework"
subtitle: "Profile-based lookup tables for international benchmarking"
author:
  - name: Lingling Ma
    affiliation: Civil Aviation University of China
  - name: Wei Dai
    affiliation: Civil Aviation University of China
  - name: Rainer Koelle
    affiliation: EUROCONTROL Performance Review Unit
date: today
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
  docx: default
  pdf: default
---

```{r}
#| label: setup
#| include: false
source(here::here("R", "setup.R"))
library(arrow)
library(dplyr)
library(ggplot2)
library(knitr)
```

```{r}
#| label: load-and-harmonize-data
#| cache: true

# Load EUR level segments
eur_segments_raw <- arrow::read_parquet(
  here::here("data-store", "derived", "eur", "level-segments-eur-2026.parquet")
)

# Load EUR harmonized milestones
eur_milestones <- arrow::read_parquet(
  here::here("data-store", "derived", "eur", "canonical-milestones-eur-2026-harmonized.parquet")
)

# ============================================================================
# Field name harmonization for current pipeline output
# ============================================================================
# NOTE: Current level-segments output uses different field names than
# expected by paper analysis. Map them here until pipeline is updated.
#
# Expected fields (for paper):        Current fields (from pipeline):
# - FLIGHT_PHASE_RAW                  -> phase_context (already derived)
# - ALT_START, ALT_END                -> start_alt, end_alt
# - altitude_band                     -> altitude_band (already computed)
#
# TODO for pipeline: Consider standardizing field names in R/level-off-milestones.R
# to match canonical milestone conventions (uppercase, explicit START/END suffixes)
# ============================================================================

eur_segments <- eur_segments_raw |>
  mutate(
    # Map phase_context to expected FLIGHT_PHASE for compatibility
    FLIGHT_PHASE = phase_context,

    # Altitude fields (keep both naming conventions)
    ALT_START = start_alt,
    ALT_END = end_alt,
    ALT_MEAN = mean_altitude_ft,

    # Time fields
    TIME_START = start_time,
    TIME_END = end_time,

    # Ensure altitude_band factor levels
    altitude_band = factor(
      altitude_band,
      levels = c("below_FL100", "FL100_FL180", "FL180_FL240", "above_FL240")
    )
  )

# Basic counts for summary
n_flights <- eur_milestones |> distinct(UID) |> nrow()
n_segments <- nrow(eur_segments)
n_paired <- eur_segments |> filter(has_end) |> nrow()
```

# Executive Summary

This document develops a refined fuel-burn estimation framework building on the 2025 ICNS CHN-EUR comparison. The key refinements address:

- **Profile characterization:** Level-segment analysis with altitude bands replaces coarse phase aggregation
- **Statistical bounds:** Percentile-based lookup tables provide upper/lower fuel burn estimates
- **Data completeness:** Implicit END marker methodology addresses systematic gaps in operational data
- **International benchmarking:** Harmonized framework supports multi-regional validation

**Current status:** EUR data complete (21,503 flights, 60,012 level segments). CHN data update in progress.

# 1. Foundation: ICNS2025 Baseline

## 1.1 Original Approach

The 2025 ICNS paper established a trajectory discretization method using characteristic milestones:

**Departure phase:**
- AOBT (actual off-block time)
- ATOT (actual take-off time)
- D100 (100NM from departure)

**Enroute phase:**
- TOC (top of climb)
- TOD (top of descent)

**Arrival phase:**
- A100 (100NM to destination)
- ALDT (actual landing time)
- AIBT (actual in-block time)

**Phase definitions:**
- TXO: AOBT → ATOT (taxi-out)
- CLIMB100: ATOT → D100 (departure climb)
- ENR100: D100 → A100 (enroute)
- DESC100: A100 → ALDT (arrival descent)
- TXI: ALDT → AIBT (taxi-in)

## 1.2 Key Findings (2025)

**Fuel burn rates (medium-type aircraft):**
- China: 35.55 kg/min average across all phases
- Europe: 34.45 kg/min average across all phases
- Overall consistency suggests comparable operational contexts

**Benefit pool estimation:**
- Average additional fuel burn: <200 kg for airborne portion
- Based on 20th percentile reference flights
- Medium-type aircraft focus

**Aircraft types studied:**
- Medium: A320, A20N (A320neo), B738 (Boeing 737-800)
- Heavy: A333 (Airbus 330-300), B789 (Boeing 787-900)

## 1.3 Limitations Identified

The 2025 approach provided valuable first results but exhibited limitations:

1. **Coarse phase characterization**
   - Single ENR100 phase conflates cruise, stepped climbs, level-offs
   - No altitude band differentiation
   - Climb/descent treated as homogeneous

2. **Profile variability unaccounted**
   - Different flight distances (500-900NM) aggregated
   - Altitude profiles not characterized
   - Level-off frequency and duration ignored

3. **Statistical approach**
   - 20th percentile reference somewhat arbitrary
   - No upper/lower bound framework
   - Limited sensitivity analysis

4. **Data quality unchecked**
   - Level marker completeness not audited
   - Orphaned segments not identified
   - QC flags absent

**Conclusion:** The milestone-based approach is sound, but operational detail is lost in aggregation. A refined framework should preserve milestones while adding level-segment intervals and profile characteristics.

# 2. Refined Methodology

## 2.1 Extended Milestone Families

The 2026 framework extends the original milestones with derived analytical events:

**Distance milestones (along-track):**
- Departure side: D040, D100, D200
- Arrival side: A200, A100, A040
- Derived from cumulative flown distance (`DIST_FLOWN_NM`)

**Flight level crossings (pressure altitude):**
- Departure side: D_FL075, D_FL100, D_FL180
- Arrival side: A_FL180, A_FL100, A_FL075
- First upward/last downward crossing logic

**Level-segment boundaries:**
- LVL_START: Beginning of level-off phase
- LVL_END: End of level-off phase
- Paired events for duration/fuel calculation

**Rationale:**
- Distance milestones enable finer phase resolution (e.g., CLIMB040 vs CLIMB100)
- FL crossings characterize vertical profile without assuming TOC/TOD
- Level segments quantify cruise efficiency directly

## 2.2 Level-Segment Detection

### 2.2.1 Source Data Challenge

PRU canonical milestone data contains LVL markers indicating level flight, but 54.4% of flights showed incomplete pairing:

| Status | Flights | % | Issue |
|--------|---------|---|-------|
| Properly paired | 8,989 | 45.6% | Complete START/END pairs |
| Odd LVL count | 10,704 | 54.4% | Missing END markers |

**Root causes identified:**
- Missing LVL_END at TOD (top of descent)
- Missing END at phase transitions (Lvl_climb → cruise)
- Trajectory data ends while aircraft still level
- Compound milestones not fully decomposed (e.g., "TOD/LVL")

### 2.2.2 Implicit END Completion

To recover usable level segments, implicit LVL_END markers are added at logical boundaries:

1. **At TOD** if preceded by active level segment
2. **At phase transitions** from Lvl_* to non-Lvl_* phase
3. **At trajectory end** if still in level phase

**Implementation:**
- Milestone multiplicity approach (separate row added)
- Same TIME/LAT/LON/ALT as boundary event
- Tagged with `.milestone_source = "implicit_end_at_TOD"`

**Example:**
```
Before completion:
  10:15:00  LVL_START  FL350
  10:45:00  TOD        FL350

After completion:
  10:15:00  LVL_START  FL350  (derived_level_segment)
  10:45:00  LVL_END    FL350  (implicit_end_at_TOD)
  10:45:00  TOD        FL350  (original)
```

**Results:**
- Paired segments: 22,236 → 54,494 (+145% usable segments)
- Orphan rate: 54.4% → 44.3%
- Remaining orphans represent true data boundaries

### 2.2.3 Quality Control Framework

All level segments tagged with QC flags:

| Flag | Count | % | Interpretation |
|------|-------|---|----------------|
| `ok` | 3,243 | 9.8% | Clean paired segment, altitude stable |
| `orphaned_no_end` | 10,704 | 32.5% | No END marker found |
| `excessive_alt_change` | 18,993 | 57.7% | >1000 ft change during "level" |

**Notes:**
- `excessive_alt_change` may indicate strict threshold (cruise-climb common)
- Flags enable sensitivity analysis by QC subset
- Duration/fuel metrics set to NA for orphaned segments

## 2.3 Profile Characterization Dimensions

The refined framework characterizes each flight by:

**1. Aircraft group:**
- Medium: A320 family, B737 family
- Heavy: A330 family, B787 family, widebodies

**2. Distance band:**
- Short: <500 NM
- Medium-short: 500-700 NM
- Medium-long: 700-1000 NM
- Long: >1000 NM

**3. Phase context:**
- Lvl_climb: Level-off during climb phase
- Lvl_cruise: Sustained cruise (implicit from TOC-TOD)
- Lvl_descent: Level-off during descent phase

**4. Altitude band:**
- below_FL100: <10,000 ft
- FL100_FL180: 10,000-18,000 ft
- FL180_FL240: 18,000-24,000 ft
- above_FL240: >24,000 ft (typical cruise)

**5. Duration class:**
- Short: <5 min
- Medium: 5-15 min
- Long: 15-30 min
- Very long: >30 min

This multidimensional characterization enables lookup tables that account for operational context rather than treating all flights as homogeneous.

## 2.4 Statistical Framework

### 2.4.1 Percentile-Based Bounds

Instead of single reference values, the framework provides percentile-based bounds:

- **P20 (efficient):** Representative of best operational practice
- **P50 (median):** Typical operational performance
- **P80 (inefficient):** Upper bound, accounts for constraints

**Rationale:**
- P20 replaces arbitrary 20% threshold from 2025
- P50-P20 spread quantifies achievable benefit pool
- P80-P50 spread indicates operational variability

### 2.4.2 Metrics per Segment

For each profile class, compute:

- **Total fuel [kg]:** Cumulative burn over interval
- **Duration [min]:** Elapsed time
- **Distance [NM]:** Flown distance
- **kg/min:** Fuel flow rate
- **kg/NM:** Distance-specific consumption
- **Level time [%]:** Proportion in level-off

### 2.4.3 Lookup Table Structure

Proposed output format:

```
aircraft_group | distance_band | phase | alt_band | metric | p20 | p50 | p80 | n
---------------|---------------|-------|----------|--------|-----|-----|-----|---
Medium         | 500-700       | cruise| >FL240   | kg/min | 32  | 35  | 39  | 450
Medium         | 500-700       | cruise| >FL240   | kg/NM  | 4.2 | 4.5 | 5.1 | 450
...
```

This structure supports:
- Profile-specific fuel burn queries
- Sensitivity analysis by percentile
- International comparison at matched profile classes
- Validation against operator data

# 3. EUR Data Exploration

## 3.1 Dataset Characteristics

**EUR Data Summary (Summer 2025):**

- **Flights:** `r formatC(n_flights, format="d", big.mark=",")` flights
- **Study airports:** 9 major European hubs (EGLL, LFPG, LTFM, LEMD, EHAM, LIRF, EDDF, EDDM, LSZH)
- **Aircraft types:** A20N, A320, A333, B738, B789
- **Level segments:** `r formatC(n_segments, format="d", big.mark=",")` total
- **Paired segments:** `r formatC(n_paired, format="d", big.mark=",")` (`r round(100*n_paired/n_segments, 1)`%)
- **Data source:** EUROCONTROL PRU canonical milestones + AEM fuel model

**Milestone counts:**

```{r}
#| label: milestone-counts
#| tbl-cap: "EUR milestone distribution"

eur_milestones |>
  count(MST, sort = TRUE) |>
  head(15) |>
  kable(col.names = c("Milestone", "Count"),
        format.args = list(big.mark = ","))
```

## 3.2 Level-Segment Duration Distributions

```{r}
#| label: duration-by-phase
#| fig-cap: "Level-segment duration by phase context"
#| fig-width: 10
#| fig-height: 6
#| warning: false

eur_segments |>
  filter(has_end, duration_sec > 0, !is.na(FLIGHT_PHASE)) |>
  mutate(duration_min = duration_sec / 60) |>
  filter(FLIGHT_PHASE %in% c("Lvl_climb", "Lvl_descent")) |>
  ggplot(aes(x = duration_min, fill = FLIGHT_PHASE)) +
  geom_histogram(bins = 50, alpha = 0.7, position = "identity") +
  facet_wrap(~FLIGHT_PHASE, ncol = 1, scales = "free_y") +
  scale_x_continuous(limits = c(0, 60), breaks = seq(0, 60, 10)) +
  labs(
    x = "Duration [minutes]",
    y = "Number of segments",
    fill = "Phase context"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

**Observations:**

```{r}
#| label: duration-summary

duration_summary <- eur_segments |>
  filter(has_end, duration_sec > 0, !is.na(FLIGHT_PHASE)) |>
  mutate(duration_min = duration_sec / 60) |>
  filter(FLIGHT_PHASE %in% c("Lvl_climb", "Lvl_descent")) |>
  group_by(FLIGHT_PHASE) |>
  summarise(
    n = n(),
    mean = mean(duration_min),
    median = median(duration_min),
    p20 = quantile(duration_min, 0.2),
    p80 = quantile(duration_min, 0.8),
    .groups = "drop"
  )

kable(duration_summary, digits = 1,
      col.names = c("Phase", "N", "Mean [min]", "Median [min]", "P20 [min]", "P80 [min]"))
```

- Lvl_climb segments tend to be shorter (median ~`r round(duration_summary$median[duration_summary$FLIGHT_PHASE == "Lvl_climb"], 0)` min)
- Lvl_descent shows longer durations (median ~`r round(duration_summary$median[duration_summary$FLIGHT_PHASE == "Lvl_descent"], 0)` min)
- Wide variability suggests operational constraints (traffic, airspace design)
- P20-P80 spread indicates benefit pool quantification potential

## 3.3 Altitude Band Analysis

```{r}
#| label: altitude-bands
#| fig-cap: "Level-segment distribution by altitude band"
#| fig-width: 10
#| fig-height: 5
#| warning: false

eur_segments |>
  filter(has_end, duration_sec > 0, !is.na(altitude_band)) |>
  count(altitude_band) |>
  ggplot(aes(x = altitude_band, y = n, fill = altitude_band)) +
  geom_col() +
  geom_text(aes(label = scales::comma(n)), vjust = -0.5) +
  labs(
    x = "Altitude band",
    y = "Number of segments",
    fill = NULL
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```

**Key findings:**

```{r}
#| label: altitude-summary

alt_summary <- eur_segments |>
  filter(has_end, duration_sec > 0, !is.na(altitude_band)) |>
  mutate(duration_min = duration_sec / 60) |>
  group_by(altitude_band) |>
  summarise(
    n = n(),
    pct = 100 * n() / sum(eur_segments$has_end & eur_segments$duration_sec > 0 & !is.na(eur_segments$altitude_band)),
    median_dur = median(duration_min),
    .groups = "drop"
  ) |>
  arrange(desc(n))

kable(alt_summary, digits = 1,
      col.names = c("Altitude band", "N", "% of total", "Median duration [min]"))
```

- Majority of level segments occur above FL240 (typical cruise altitude)
- FL180-FL240 shows substantial activity (stepped climbs, traffic constraints)
- Below FL100 segments rare but operationally significant (holding, sequencing)
- Altitude-specific fuel burn rates essential for accurate estimation

## 3.4 Aircraft Type Patterns

```{r}
#| label: type-analysis
#| fig-cap: "Level-segment duration by aircraft type"
#| fig-width: 10
#| fig-height: 6
#| warning: false

eur_segments |>
  filter(has_end, duration_sec > 0, !is.na(TYPE)) |>
  filter(TYPE %in% c("A320", "A20N", "B738", "A333", "B789")) |>
  mutate(
    duration_min = duration_sec / 60,
    aircraft_group = ifelse(TYPE %in% c("A320", "A20N", "B738"),
                            "Medium", "Heavy")
  ) |>
  ggplot(aes(x = TYPE, y = duration_min, fill = aircraft_group)) +
  geom_boxplot() +
  coord_cartesian(ylim = c(0, 40)) +
  labs(
    x = "Aircraft type",
    y = "Duration [minutes]",
    fill = "Aircraft group"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

**Aircraft-specific observations:**

```{r}
#| label: type-summary

type_summary <- eur_segments |>
  filter(has_end, duration_sec > 0, !is.na(TYPE)) |>
  filter(TYPE %in% c("A320", "A20N", "B738", "A333", "B789")) |>
  mutate(duration_min = duration_sec / 60) |>
  group_by(TYPE) |>
  summarise(
    n = n(),
    median_dur = median(duration_min),
    p20_dur = quantile(duration_min, 0.2),
    p80_dur = quantile(duration_min, 0.8),
    .groups = "drop"
  ) |>
  arrange(desc(n))

kable(type_summary, digits = 1,
      col.names = c("Type", "N", "Median [min]", "P20 [min]", "P80 [min]"))
```

- Medium types show similar level-segment patterns (operational consistency)
- Heavy types exhibit wider duration variability (longer flights, different constraints)
- Type-specific lookup tables required for accurate fuel burn estimation

## 3.5 Quality Control Insights

```{r}
#| label: qc-flags
#| tbl-cap: "Level-segment QC flag distribution"

eur_segments |>
  count(qc_flag, sort = TRUE) |>
  mutate(pct = 100 * n / sum(n)) |>
  kable(digits = 1,
        col.names = c("QC Flag", "Count", "% of total"),
        format.args = list(big.mark = ","))
```

**QC implications:**

- `excessive_alt_change` (57.7%): Current 1000 ft threshold may be too strict
  - Cruise-climb operations common in European airspace
  - Stepped climbs for efficiency/traffic
  - Threshold review recommended (1500-2000 ft more realistic)

- `orphaned_no_end` (32.5%): Represents true data boundaries
  - Implicit END completion reduced from 54.4%
  - Remaining orphans at trajectory edges or missing TOD
  - Acceptable for frequency analysis, excluded from duration/fuel

- `ok` (9.8%): Conservative clean segment subset
  - Useful for validation and sensitivity analysis
  - May exclude operationally valid segments

**Recommendation:** Primary analysis should use `has_end = TRUE` (includes excessive_alt_change), with `ok`-only sensitivity check.

# 4. Lookup Table Framework

## 4.1 Proposed Structure

The refined lookup table provides fuel burn estimates stratified by operational context:

**Dimensions:**
1. **Aircraft group:** Medium (A320/B737 family), Heavy (widebodies)
2. **Distance band:** <500, 500-700, 700-1000, >1000 NM
3. **Phase:** TXO, CLIMB, ENR, DESC, TXI
4. **Altitude band:** <FL100, FL100-FL180, FL180-FL240, >FL240
5. **Profile class:** Short-level, Medium-level, Long-level, No-level

**Metrics per cell:**
- Fuel burn [kg]: P20, P50, P80
- Duration [min]: P20, P50, P80
- kg/min: P20, P50, P80
- kg/NM: P20, P50, P80
- Sample size: n

## 4.2 Example: EUR Medium Types, Cruise Level

```{r}
#| label: example-lookup
#| tbl-cap: "Example lookup table entry (EUR, Medium, >FL240)"
#| warning: false

# This is a placeholder - full implementation pending fuel data integration
# Showing structure with duration data available

eur_segments |>
  filter(has_end, duration_sec > 0, !is.na(TYPE), !is.na(altitude_band)) |>
  filter(TYPE %in% c("A320", "A20N", "B738")) |>
  filter(altitude_band == "above_FL240") |>
  mutate(duration_min = duration_sec / 60) |>
  summarise(
    n = n(),
    dur_p20 = quantile(duration_min, 0.2),
    dur_p50 = quantile(duration_min, 0.5),
    dur_p80 = quantile(duration_min, 0.8),
    .groups = "drop"
  ) |>
  kable(digits = 1,
        col.names = c("N", "Duration P20 [min]", "P50 [min]", "P80 [min]"))
```

**Interpretation:**
- P20 duration represents efficient operations (shortest level-segment durations)
- P50-P20 spread quantifies typical operational variability
- P80 represents constrained operations (holding, traffic sequencing)

**Next step:** Integrate fuel burn data (fuel_burn_kg field) to populate kg/min and kg/NM metrics.

## 4.3 Validation Strategy

**Internal validation:**
1. Compare P50 values to ICNS2025 aggregated rates (35 kg/min baseline)
2. Check consistency across adjacent distance bands
3. Verify altitude band trends (higher FL = lower kg/NM typically)

**External validation:**
- Operator fuel burn data (confidential, aggregated comparison)
- BADA/AEM model cross-checks
- openAP model comparisons (open-source benchmark)

**Sensitivity analysis:**
- QC flag subsets (ok-only vs has_end)
- Altitude threshold variations (1000 ft vs 1500 ft)
- Duration filters (exclude <5 min segments)

## 4.4 CHN-EUR Comparison Framework

Once CHN data available, comparison approach:

1. **Profile matching:** Same distance band, aircraft group, altitude
2. **Percentile comparison:** Compare P20/P50/P80 across regions
3. **Operational context:** Account for airspace design differences
4. **Statistical testing:** Significance of regional differences

**Hypothesis:** Medium-type aircraft show similar fuel burn rates when profile-matched (ICNS2025 suggested 35 kg/min convergence). Refined framework should confirm at finer granularity.

# 5. CHN Integration Strategy

## 5.1 Expected CHN Data Structure

**Source:** QAR/FDR data from CAAC-OSC operations

**Expected fields:**
- High-frequency trajectory: TIME, LAT, LON, ALT
- Fuel flow per engine (FFxC fields)
- Aircraft metadata: TYPE, ADEP, ADES
- Cumulative fuel burn (derived from fuel flow)

**Milestones expected:**
- Operational: AOBT, ATOT, ALDT, AIBT
- Derived: D040, D100, D200, A200, A100, A040
- FL crossings: D_FL075, D_FL100, D_FL180, A_FL180, A_FL100, A_FL075
- Level segments: LVL_START, LVL_END (if derivable)

## 5.2 Harmonization Requirements

**Coordinate with CHN colleagues on:**

1. **Fuel flow units:** Confirm kg/h vs lb/h (QC framework established)
2. **Block times:** First/last trajectory point = AOBT/AIBT? Or engine on/off?
3. **Milestone derivation:** Same distance/FL crossing logic as EUR
4. **Level-segment detection:** Can LVL_START/END be derived from QAR altitude stability?

**Data processing pipeline:**
- Apply same canonical milestone functions (R/chn-data-prep.R exists)
- Generate level-segment intervals with QC flags
- Produce comparable profile characterization tables

## 5.3 Comparison Dimensions

**Matched profile comparison:**

| Dimension | EUR Data | CHN Data | Match Criteria |
|-----------|----------|----------|----------------|
| Aircraft | A320, B738 | A320, B738 | Same ICAO type |
| Distance | 500-700 NM | 500-700 NM | ±50 NM tolerance |
| Altitude | >FL240 | >FL240 | Same band |
| Phase | Cruise level | Cruise level | Same context |

**Expected outputs:**
- Side-by-side percentile tables
- Regional difference quantification
- Operational context interpretation (airspace design, traffic density)

## 5.4 Placeholders for CHN Results

**[To be populated when CHN data available]**

- CHN dataset characteristics (flights, segments, airports)
- CHN level-segment duration distributions
- CHN altitude band patterns
- CHN-EUR profile-matched comparison tables
- Convergence analysis (35 kg/min hypothesis test at finer granularity)

# 6. Discussion: Towards Operational Lookup Tables

## 6.1 Key Advances Over ICNS2025

**Methodological improvements:**

1. **Profile characterization:** Altitude bands and level-segment analysis replace coarse aggregation
   - Enables fuel burn estimation sensitive to operational context
   - Accounts for stepped climbs, cruise-climb, traffic constraints

2. **Data completeness:** Implicit END methodology recovers 10% of orphaned segments
   - 54.4% → 44.3% orphan rate improvement
   - Transparent QC flags enable sensitivity analysis
   - Remaining orphans represent true data boundaries

3. **Statistical framework:** Percentile-based bounds replace single reference
   - P20/P50/P80 quantify operational variability and benefit pools
   - Enables "what-if" analysis for ANS interventions
   - Supports uncertainty quantification

4. **Reproducibility:** R2 cloud storage with manifest tracking
   - All data products versioned and accessible
   - Processing pipeline documented (technical-note-data-preparation.qmd)
   - Code modular and reusable (R/ function modules)

## 6.2 Data Quality Lessons

**Operational data characteristics:**

- **Level marker incompleteness is systematic, not random**
  - 54.4% of EUR flights had odd LVL counts (missing ENDs)
  - Root causes: missing markers at TOD, phase transitions, trajectory boundaries
  - Solution: Implicit END completion at logical boundaries

- **Altitude stability thresholds need regional calibration**
  - 57.7% flagged for excessive_alt_change (>1000 ft)
  - European airspace: cruise-climb common for efficiency
  - Threshold review: 1500-2000 ft more appropriate

- **Phase label consistency varies by data source**
  - EUR: Lvl_climb, Lvl_descent explicitly labeled
  - CHN: May need altitude-based phase inference from QAR
  - Harmonization essential for valid comparison

**Recommendation for PRU colleagues:**
- Document LVL pairing status in data dictionary
- Review LVL derivation algorithm to add END markers at TOD/phase transitions
- Provide per-release pairing statistics

## 6.3 International Benchmarking Implications

**Harmonization requirements:**

1. **Milestone definitions:** Aligned distance/FL crossing logic documented
2. **Level-segment criteria:** Consistent altitude stability thresholds
3. **QC framework:** Common flag taxonomy (ok, orphaned, excessive_alt_change)
4. **Profile dimensions:** Agreed aircraft groups, distance bands, altitude bands

**Multi-regional validation potential:**

- EUR + CHN baseline (current study)
- Extension to other ICAO regions (NAM, ASIA/PAC, MID)
- Contribution to ICAO GANP KPI16 guidance (additional fuel burn)
- Support for regional performance comparison frameworks

**GANP KPI16 enhancement:**

Current KPI16 converts additional time/distance to fuel burn using average rates. Refined lookup tables enable:

- **Context-specific conversion:** Use profile-matched rates instead of global average
- **Benefit pool quantification:** P50-P20 spread estimates achievable savings
- **Sensitivity bounds:** P80-P20 range for uncertainty
- **Regional comparability:** Harmonized methodology across ANSPs

## 6.4 Limitations and Future Work

**Current limitations:**

1. **Fuel data integration incomplete**
   - EUR: fuel_burn_kg field present but not yet integrated into analysis
   - Need to validate fuel burn calculations per level segment
   - Cross-check against AEM model outputs

2. **CHN data pending**
   - Updated CHN data processing in progress
   - Comparison results awaiting harmonized CHN level segments
   - Fuel flow unit confirmation needed (kg/h vs lb/h)

3. **Sample representativeness**
   - EUR: Single month (Summer 2025), 9 airports
   - CHN: Previous sample ~100 flights, limited airports
   - Seasonal variability not assessed
   - Need multi-year validation

4. **Profile class granularity**
   - Current: distance band + altitude + phase context
   - Future: Add weather (wind, temperature), time-of-day, traffic density
   - Machine learning potential for profile clustering

**Next steps:**

**Short-term (Q4 2026):**
- Integrate EUR fuel burn data (fuel_burn_kg → kg/min, kg/NM metrics)
- Receive and process updated CHN data
- Generate CHN-EUR profile-matched comparison tables
- Validate P50 values against ICNS2025 baseline (35 kg/min)

**Medium-term (2027):**
- Extend EUR data to multi-year sample (seasonal variability)
- Expand aircraft type coverage (additional medium/heavy types)
- Publish lookup table methodology paper
- Contribute to ICAO GANP KPI16 guidance update

**Long-term (2027+):**
- Multi-regional validation (NAM, ASIA/PAC, MID regions)
- Operational implementation: ANSP fuel burn monitoring dashboards
- Machine learning profile clustering
- Integration with emissions inventories (non-CO₂ effects)

## 6.5 Pipeline Field Name Standardization

**Current field name mappings applied in this document:**

The EUR level-segment pipeline (`R/level-off-milestones.R`) outputs fields with lowercase/mixed naming:
- `phase_context` (mapped to `FLIGHT_PHASE` for paper)
- `start_alt`, `end_alt` (mapped to `ALT_START`, `ALT_END`)
- `start_time`, `end_time` (mapped to `TIME_START`, `TIME_END`)

**Recommendation for pipeline workers:**

Consider standardizing field names in `R/level-off-milestones.R` to match canonical milestone conventions:
- Uppercase base names: `PHASE`, `ALT`, `TIME`, `FUEL`
- Explicit START/END suffixes: `PHASE_START`, `ALT_START`, `TIME_START`
- Derived fields prefixed: `ALT_MEAN`, `ALT_CHANGE`, `FUEL_BURN`

This would eliminate the need for field name mapping in analysis documents and improve consistency across the processing pipeline.

# 7. Conclusions

This exploratory document establishes a refined fuel-burn estimation framework addressing limitations of the 2025 ICNS baseline:

**Methodological advances:**
- Level-segment analysis with altitude bands (60,012 EUR segments characterized)
- Implicit END completion methodology (orphan rate 54.4% → 44.3%)
- Percentile-based lookup tables (P20/P50/P80 bounds)
- Comprehensive QC framework for data quality transparency

**EUR data insights:**
- Majority of level segments above FL240 (typical cruise)
- Median durations vary by phase context (climb shorter than descent)
- Wide operational variability (P20-P80 spread) quantifies benefit pools
- Medium types show consistent patterns (A320/B738)

**Framework status:**
- EUR data complete and analyzed (Summer 2025, 21,503 flights)
- CHN data integration in progress
- Lookup table structure defined, pending fuel data integration
- Reproducible pipeline with R2 cloud storage

**International benchmarking readiness:**
- Harmonized milestone definitions documented
- QC framework supports multi-regional validation
- GANP KPI16 enhancement potential demonstrated
- Path to operational lookup tables established

**The refined framework provides a foundation for profile-aware fuel burn estimation supporting ICAO net-zero 2050 goals through transparent, reproducible, and internationally comparable performance monitoring.**

---

# References

1. ICAO (2022). Long-Term Aspirational Goal, Special Supplement. [https://www.icao.int/environmental-protection/Documents/EnvironmentalReports/2022/ENVReport2022_Special%20Supplement%20on%20LTAG.pdf](https://www.icao.int/environmental-protection/Documents/EnvironmentalReports/2022/ENVReport2022_Special%20Supplement%20on%20LTAG.pdf)

2. ICAO (2023). GANP KPI Overview. [https://www4.icao.int/ganpportal/ASBU/KPI](https://www4.icao.int/ganpportal/ASBU/KPI)

3. Ma, L., Dai, W., & Koelle, R. (2025). Fuel flow analysis for different flight phases based on a data-driven approach. ICNS 2025 Conference Proceedings.

4. Nuic, A. (2009). Base of Aircraft Data (BADA) Product Management Document. EUROCONTROL.

5. Deransy, R. (2022). EUROCONTROL - AEM, Advanced Emission Model, Flight Emissions Assessment. [https://www.eurocontrol.int/sites/default/files/2022-12/ectl_env_impact_assessment_forum_-_aem.pdf](https://www.eurocontrol.int/sites/default/files/2022-12/ectl_env_impact_assessment_forum_-_aem.pdf)

6. Sun, J., Hoekstra, J., & Ellerbroek, J. (2020). OpenAP: An open-source aircraft performance model for air transportation studies and simulations. *Aerospace*, 7(8), 104. doi:10.3390/aerospace7080104

---

**Document Status:** Exploratory analysis (October 2026)
**EUR Data:** Complete
**CHN Data:** Pending update
**Next Update:** Upon CHN data integration
