---
title: "Refined Fuel Burn Estimation Framework"
subtitle: "Profile-based lookup tables for international benchmarking"
author:
  - name: Lingling Ma
    affiliation: Civil Aviation University of China
  - name: Wei Dai
    affiliation: Civil Aviation University of China
  - name: Rainer Koelle
    affiliation: EUROCONTROL Performance Review Unit
date: today
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
  docx: default
  pdf: default
---

```{r}
#| label: setup
#| include: false
source(here::here("R", "setup.R"))
library(arrow)
library(dplyr)
library(ggplot2)
library(knitr)
library(tidyr)
```

```{r}
#| label: load-and-harmonize-data
#| cache: true

data_store <- Sys.getenv("FUELBURN_DATA_STORE", unset = here::here(".fuelburn-data-store"))

processed_path <- function(...) {
  file.path(data_store, ...)
}

eur_milestones <- arrow::read_parquet(
  processed_path("processed", "eur", "EUR-canonical-milestones-summer2025.parquet")
)
eur_segments_raw <- arrow::read_parquet(
  processed_path("processed", "eur", "EUR-level-segments-summer2025.parquet")
)
chn_milestones <- arrow::read_parquet(
  processed_path("processed", "chn", "CHN-canonical-milestones-summer2025.parquet")
)
chn_segments_raw <- arrow::read_parquet(
  processed_path("processed", "chn", "CHN-level-segments-summer2025.parquet")
)
chn_phase_summaries <- arrow::read_parquet(
  processed_path("processed", "chn", "CHN-phase-summaries-summer2025.parquet")
)

altitude_band_from_ft <- function(x) {
  case_when(
    is.na(x) ~ NA_character_,
    x < 10000 ~ "below_FL100",
    x < 18000 ~ "FL100_FL180",
    x < 24000 ~ "FL180_FL240",
    TRUE ~ "above_FL240"
  )
}

standardise_level_segments <- function(data, region) {
  data |>
    mutate(
      REGION = region,
      HAS_END = if ("HAS_END" %in% names(data)) .data$HAS_END else !is.na(.data$END_TIME),
      FUEL_BURN_KG = if ("FUEL_BURN_KG" %in% names(data)) .data$FUEL_BURN_KG else .data$FUEL_KG,
      MEAN_ALTITUDE_FT = if ("MEAN_ALTITUDE_FT" %in% names(data)) {
        .data$MEAN_ALTITUDE_FT
      } else {
        (.data$START_ALT_FT + .data$END_ALT_FT) / 2
      },
      ALTITUDE_BAND = if ("ALTITUDE_BAND" %in% names(data)) {
        .data$ALTITUDE_BAND
      } else {
        altitude_band_from_ft(.data$MEAN_ALTITUDE_FT)
      },
      LEVEL_CONTEXT_PHASE = if ("LEVEL_CONTEXT_PHASE" %in% names(data)) {
        .data$LEVEL_CONTEXT_PHASE
      } else {
        .data$PHASE_CONTEXT
      },
      qc_flag = if ("qc_flag" %in% names(data)) .data$qc_flag else "ok",
      duration_min = .data$DURATION_SEC / 60,
      kg_min = .data$FUEL_BURN_KG / .data$duration_min,
      altitude_band = factor(
        .data$ALTITUDE_BAND,
        levels = c("below_FL100", "FL100_FL180", "FL180_FL240", "above_FL240")
      )
    )
}

eur_segments <- standardise_level_segments(eur_segments_raw, "EUR")
chn_segments <- standardise_level_segments(chn_segments_raw, "CHN")
level_segments <- bind_rows(eur_segments, chn_segments)

milestones <- bind_rows(
  eur_milestones |> mutate(REGION = "EUR"),
  chn_milestones |> mutate(REGION = "CHN")
)

dataset_summary <- tibble::tibble(
  REGION = c("EUR", "CHN"),
  flights = c(n_distinct(eur_milestones$SOURCE_UID), n_distinct(chn_milestones$SOURCE_UID)),
  milestones = c(nrow(eur_milestones), nrow(chn_milestones)),
  level_segments = c(nrow(eur_segments), nrow(chn_segments)),
  clean_segments = c(sum(eur_segments$qc_flag == "ok", na.rm = TRUE), sum(chn_segments$qc_flag == "ok", na.rm = TRUE))
) |>
  mutate(clean_share = clean_segments / level_segments)

eur_flights <- dataset_summary$flights[dataset_summary$REGION == "EUR"]
chn_flights <- dataset_summary$flights[dataset_summary$REGION == "CHN"]
eur_n_segments <- dataset_summary$level_segments[dataset_summary$REGION == "EUR"]
chn_n_segments <- dataset_summary$level_segments[dataset_summary$REGION == "CHN"]
```

# Executive Summary

This document develops a refined fuel-burn estimation framework building on the 2025 ICNS CHN-EUR comparison. The key refinements address:

- **Profile characterization:** Level-segment analysis with altitude bands replaces coarse phase aggregation
- **Statistical bounds:** Percentile-based lookup tables provide upper/lower fuel burn estimates
- **Data completeness:** The FIXED European pipeline and the new Chinese QAR products provide complete level-segment interval tables for first comparative exploration
- **International benchmarking:** Harmonized framework supports multi-regional validation

**Current status:** EUR and CHN summer 2025 production products are available through the R2 manifest. The exploratory baseline now covers `r formatC(eur_flights, format="d", big.mark=",")` European flights with `r formatC(eur_n_segments, format="d", big.mark=",")` level segments and `r formatC(chn_flights, format="d", big.mark=",")` Chinese flights with `r formatC(chn_n_segments, format="d", big.mark=",")` level segments.

# 1. Foundation: ICNS2025 Baseline

## 1.1 Original Approach

The 2025 ICNS paper established a trajectory discretization method using characteristic milestones:

**Departure phase:**
- AOBT (actual off-block time)
- ATOT (actual take-off time)
- D100 (100NM from departure)

**Enroute phase:**
- TOC (top of climb)
- TOD (top of descent)

**Arrival phase:**
- A100 (100NM to destination)
- ALDT (actual landing time)
- AIBT (actual in-block time)

**Phase definitions:**
- TXO: AOBT → ATOT (taxi-out)
- CLIMB100: ATOT → D100 (departure climb)
- ENR100: D100 → A100 (enroute)
- DESC100: A100 → ALDT (arrival descent)
- TXI: ALDT → AIBT (taxi-in)

## 1.2 Key Findings (2025)

**Fuel burn rates (medium-type aircraft):**
- China: 35.55 kg/min average across all phases
- Europe: 34.45 kg/min average across all phases
- Overall consistency suggests comparable operational contexts

**Benefit pool estimation:**
- Average additional fuel burn: <200 kg for airborne portion
- Based on 20th percentile reference flights
- Medium-type aircraft focus

**Aircraft types studied:**
- Medium: A320, A20N (A320neo), B738 (Boeing 737-800)
- Heavy: A333 (Airbus 330-300), B789 (Boeing 787-900)

## 1.3 Limitations Identified

The 2025 approach provided valuable first results but exhibited limitations:

1. **Coarse phase characterization**
   - Single ENR100 phase conflates cruise, stepped climbs, level-offs
   - No altitude band differentiation
   - Climb/descent treated as homogeneous

2. **Profile variability unaccounted**
   - Different flight distances (500-900NM) aggregated
   - Altitude profiles not characterized
   - Level-off frequency and duration ignored

3. **Statistical approach**
   - 20th percentile reference somewhat arbitrary
   - No upper/lower bound framework
   - Limited sensitivity analysis

4. **Data quality unchecked**
   - Level marker completeness not audited
   - Orphaned segments not identified
   - QC flags absent

**Conclusion:** The milestone-based approach is sound, but operational detail is lost in aggregation. A refined framework should preserve milestones while adding level-segment intervals and profile characteristics.

# 2. Refined Methodology

## 2.1 Extended Milestone Families

The 2026 framework extends the original milestones with derived analytical events:

**Distance milestones (along-track):**
- Departure side: D040, D100, D200
- Arrival side: A200, A100, A040
- Derived from cumulative flown distance (`DIST_FLOWN_NM`)

**Flight level crossings (pressure altitude):**
- Departure side: D_FL075, D_FL100, D_FL180
- Arrival side: A_FL180, A_FL100, A_FL075
- First upward/last downward crossing logic

**Level-segment boundaries:**
- LVL_START: Beginning of level-off phase
- LVL_END: End of level-off phase
- Paired events for duration/fuel calculation

**Rationale:**
- Distance milestones enable finer phase resolution (e.g., CLIMB040 vs CLIMB100)
- FL crossings characterize vertical profile without assuming TOC/TOD
- Level segments quantify cruise efficiency directly

## 2.2 Level-Segment Detection

### 2.2.1 Source-Specific Extraction

The exploratory paper now starts from the production data products published through the R2 manifest rather than from intermediate EUR-only development files.

For Europe, the FIXED PRU pipeline corrects the earlier LOBT filter issue, preserves single-role milestone rows, and writes a production level-segment table with explicit start/end intervals. For China, the QAR-derived products provide canonical milestones, level-segment intervals, and phase summaries for the selected summer 2025 sample.

The two source paths are deliberately kept source-specific up to the processed products:

- EUR/PRU: surveillance-derived G2G extraction, AEM fuel model, enriched canonical milestones, and FIXED level intervals.
- CHN/QAR: aircraft-recorded trajectory and fuel-flow processing, canonical milestone generation, and QAR-derived level intervals.

### 2.2.2 Harmonised Analysis Layer

The common analysis layer keeps a small shared set of level-segment fields:

- `REGION`, `SOURCE_UID`, `ADEP`, `ADES`, and `TYPE`;
- start/end time, altitude, fuel, and duration;
- `LEVEL_CONTEXT_PHASE` for broad climb/enroute/descent context;
- `ALTITUDE_BAND` for vertical profile stratification;
- source-level QC flags where available.

This lets the paper compare first-order profile descriptors without pretending that the PRU and QAR extraction mechanics are identical.

### 2.2.3 Quality Control Framework

For this exploratory baseline, quality flags are retained rather than filtered away. The EUR production table marks segments with large altitude changes during a nominal level interval, while the CHN level-segment table currently enters the common layer as `ok`.

```{r}
#| label: level-qc-overview
#| tbl-cap: "Level-segment QC distribution in the processed EUR/CHN products"

level_segments |>
  count(REGION, qc_flag, name = "segments") |>
  group_by(REGION) |>
  mutate(pct = 100 * segments / sum(segments)) |>
  ungroup() |>
  arrange(REGION, desc(segments)) |>
  kable(
    digits = 1,
    col.names = c("Region", "QC flag", "Segments", "% within region"),
    format.args = list(big.mark = ",")
  )
```

The immediate analysis should therefore report both all-segment and clean-segment sensitivities. The next modelling step should decide whether altitude-change flags represent invalid level intervals or operationally meaningful cruise-climb/step-climb behaviour.

## 2.3 Profile Characterization Dimensions

The refined framework characterizes each flight by:

**1. Aircraft group:**
- Medium: A320 family, B737 family
- Heavy: A330 family, B787 family, widebodies

**2. Distance band:**
- Short: <500 NM
- Medium-short: 500-700 NM
- Medium-long: 700-1000 NM
- Long: >1000 NM

**3. Phase context:**
- CLIMB: Level-off during climb context
- ENROUTE: Sustained enroute/cruise level context
- DESCENT: Level-off during descent context

**4. Altitude band:**
- below_FL100: <10,000 ft
- FL100_FL180: 10,000-18,000 ft
- FL180_FL240: 18,000-24,000 ft
- above_FL240: >24,000 ft (typical cruise)

**5. Duration class:**
- Short: <5 min
- Medium: 5-15 min
- Long: 15-30 min
- Very long: >30 min

This multidimensional characterization enables lookup tables that account for operational context rather than treating all flights as homogeneous.

## 2.4 Statistical Framework

### 2.4.1 Percentile-Based Bounds

Instead of single reference values, the framework provides percentile-based bounds:

- **P20 (efficient):** Representative of best operational practice
- **P50 (median):** Typical operational performance
- **P80 (inefficient):** Upper bound, accounts for constraints

**Rationale:**
- P20 replaces arbitrary 20% threshold from 2025
- P50-P20 spread quantifies achievable benefit pool
- P80-P50 spread indicates operational variability

### 2.4.2 Metrics per Segment

For each profile class, compute:

- **Total fuel [kg]:** Cumulative burn over interval
- **Duration [min]:** Elapsed time
- **Distance [NM]:** Flown distance
- **kg/min:** Fuel flow rate
- **kg/NM:** Distance-specific consumption
- **Level time [%]:** Proportion in level-off

### 2.4.3 Lookup Table Structure

Proposed output format:

```
aircraft_group | distance_band | phase | alt_band | metric | p20 | p50 | p80 | n
---------------|---------------|-------|----------|--------|-----|-----|-----|---
Medium         | 500-700       | cruise| >FL240   | kg/min | 32  | 35  | 39  | 450
Medium         | 500-700       | cruise| >FL240   | kg/NM  | 4.2 | 4.5 | 5.1 | 450
...
```

This structure supports:
- Profile-specific fuel burn queries
- Sensitivity analysis by percentile
- International comparison at matched profile classes
- Validation against operator data

# 3. EUR and CHN Data Exploration

## 3.1 Dataset Characteristics

**Processed Data Summary (Summer 2025):**

- **EUR:** `r formatC(eur_flights, format="d", big.mark=",")` flights, `r formatC(nrow(eur_milestones), format="d", big.mark=",")` canonical milestone rows, `r formatC(eur_n_segments, format="d", big.mark=",")` level segments.
- **CHN:** `r formatC(chn_flights, format="d", big.mark=",")` flights, `r formatC(nrow(chn_milestones), format="d", big.mark=",")` canonical milestone rows, `r formatC(chn_n_segments, format="d", big.mark=",")` level segments.
- **Data source:** R2 manifest production products under `processed/eur` and `processed/chn`.
- **Purpose:** establish common exploratory descriptors before building fuel-burn lookup tables.

```{r}
#| label: dataset-summary
#| tbl-cap: "Processed production data products available for exploratory analysis"

dataset_summary |>
  mutate(clean_share = 100 * clean_share) |>
  kable(
    digits = 1,
    col.names = c("Region", "Flights", "Milestone rows", "Level segments", "Clean segments", "Clean share [%]"),
    format.args = list(big.mark = ",")
  )
```

**Milestone counts by region:**

```{r}
#| label: milestone-counts
#| tbl-cap: "Top canonical milestone labels by region"

milestones |>
  count(REGION, MST, sort = TRUE) |>
  group_by(REGION) |>
  slice_max(n, n = 12, with_ties = FALSE) |>
  ungroup() |>
  arrange(REGION, desc(n)) |>
  kable(
    col.names = c("Region", "Milestone", "Count"),
    format.args = list(big.mark = ",")
  )
```

## 3.2 Level-Segment Duration Distributions

```{r}
#| label: duration-by-phase
#| fig-cap: "Level-segment duration by region and phase context"
#| fig-width: 10
#| fig-height: 6
#| warning: false

level_segments |>
  filter(HAS_END, DURATION_SEC > 0, !is.na(LEVEL_CONTEXT_PHASE)) |>
  filter(LEVEL_CONTEXT_PHASE %in% c("CLIMB", "ENROUTE", "DESCENT")) |>
  ggplot(aes(x = duration_min, fill = REGION)) +
  geom_histogram(bins = 50, alpha = 0.7, position = "identity") +
  facet_grid(REGION ~ LEVEL_CONTEXT_PHASE, scales = "free_y") +
  scale_x_continuous(limits = c(0, 60), breaks = seq(0, 60, 10)) +
  labs(
    x = "Duration [minutes]",
    y = "Number of segments",
    fill = "Region"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

**Observations:**

```{r}
#| label: duration-summary

duration_summary <- eur_segments |>
  filter(HAS_END, DURATION_SEC > 0, !is.na(LEVEL_CONTEXT_PHASE)) |>
  mutate(REGION = "EUR") |>
  bind_rows(
    chn_segments |>
      filter(HAS_END, DURATION_SEC > 0, !is.na(LEVEL_CONTEXT_PHASE)) |>
      mutate(REGION = "CHN")
  ) |>
  filter(LEVEL_CONTEXT_PHASE %in% c("CLIMB", "ENROUTE", "DESCENT")) |>
  group_by(REGION, LEVEL_CONTEXT_PHASE) |>
  summarise(
    n = n(),
    mean = mean(duration_min),
    median = median(duration_min),
    p20 = quantile(duration_min, 0.2),
    p80 = quantile(duration_min, 0.8),
    .groups = "drop"
  )

kable(duration_summary, digits = 1,
      col.names = c("Region", "Phase context", "N", "Mean [min]", "Median [min]", "P20 [min]", "P80 [min]"),
      format.args = list(big.mark = ","))
```

The first comparison shows both source and regional effects. CHN contains many short climb/descent level intervals and longer enroute intervals, while EUR has a larger share of long enroute segments and a wider spread in flagged altitude-changing intervals. These patterns are descriptive at this stage; they should be interpreted together with extraction rules and QC flags.

## 3.3 Altitude Band Analysis

```{r}
#| label: altitude-bands
#| fig-cap: "Level-segment distribution by altitude band"
#| fig-width: 10
#| fig-height: 5
#| warning: false

level_segments |>
  filter(HAS_END, DURATION_SEC > 0, !is.na(altitude_band)) |>
  count(REGION, altitude_band) |>
  ggplot(aes(x = altitude_band, y = n, fill = REGION)) +
  geom_col(position = "dodge") +
  geom_text(aes(label = scales::comma(n)), position = position_dodge(width = 0.9), vjust = -0.4, size = 3) +
  labs(
    x = "Altitude band",
    y = "Number of segments",
    fill = "Region"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

**Key findings:**

```{r}
#| label: altitude-summary

alt_summary <- level_segments |>
  filter(HAS_END, DURATION_SEC > 0, !is.na(altitude_band)) |>
  group_by(REGION, altitude_band) |>
  summarise(
    n = n(),
    pct = 100 * n() / sum(level_segments$REGION == first(REGION) & level_segments$HAS_END & level_segments$DURATION_SEC > 0 & !is.na(level_segments$altitude_band)),
    median_dur = median(duration_min),
    .groups = "drop"
  ) |>
  arrange(REGION, altitude_band)

kable(alt_summary, digits = 1,
      col.names = c("Region", "Altitude band", "N", "% within region", "Median duration [min]"),
      format.args = list(big.mark = ","))
```

- Above-FL240 intervals dominate both regions, but both datasets also contain substantial lower-altitude level intervals.
- Below-FL100 and FL100-FL180 segments are operationally important for sequencing, constrained climbs, and arrival management.
- Altitude-specific fuel burn rates essential for accurate estimation

## 3.4 Aircraft Type Patterns

```{r}
#| label: type-analysis
#| fig-cap: "Level-segment duration by aircraft type"
#| fig-width: 10
#| fig-height: 6
#| warning: false

level_segments |>
  filter(HAS_END, DURATION_SEC > 0, !is.na(TYPE)) |>
  mutate(
    aircraft_group = case_when(
      str_detect(TYPE, "A319|A320|A321|A20N|A21N|B737|B738|B739|C919") ~ "Medium",
      TRUE ~ "Heavy/Other"
    )
  ) |>
  group_by(REGION, TYPE) |>
  filter(n() >= 250) |>
  ungroup() |>
  ggplot(aes(x = reorder(TYPE, duration_min, median), y = duration_min, fill = REGION)) +
  geom_boxplot() +
  coord_cartesian(ylim = c(0, 40)) +
  facet_wrap(~REGION, scales = "free_x") +
  labs(
    x = "Aircraft type",
    y = "Duration [minutes]",
    fill = "Region"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom", axis.text.x = element_text(angle = 45, hjust = 1))
```

**Aircraft-specific observations:**

```{r}
#| label: type-summary

type_summary <- level_segments |>
  filter(HAS_END, DURATION_SEC > 0, !is.na(TYPE)) |>
  group_by(REGION, TYPE) |>
  summarise(
    n = n(),
    median_dur = median(duration_min),
    p20_dur = quantile(duration_min, 0.2),
    p80_dur = quantile(duration_min, 0.8),
    .groups = "drop"
  ) |>
  group_by(REGION) |>
  slice_max(n, n = 8, with_ties = FALSE) |>
  ungroup() |>
  arrange(REGION, desc(n))

kable(type_summary, digits = 1,
      col.names = c("Region", "Type", "N", "Median [min]", "P20 [min]", "P80 [min]"),
      format.args = list(big.mark = ","))
```

- The common type set is not identical across the two samples, so type-specific lookup tables need a mapping layer for aircraft family/class.
- The Chinese sample includes additional domestic and long-haul types; the European sample is dominated by short/medium-haul narrowbodies.
- Type-specific stratification remains necessary before fuel-rate comparisons.

## 3.5 Quality Control Insights

```{r}
#| label: qc-flags
#| tbl-cap: "Level-segment QC flag distribution"

level_segments |>
  count(REGION, qc_flag, sort = TRUE) |>
  group_by(REGION) |>
  mutate(pct = 100 * n / sum(n)) |>
  ungroup() |>
  arrange(REGION, desc(n)) |>
  kable(digits = 1,
        col.names = c("Region", "QC Flag", "Count", "% within region"),
        format.args = list(big.mark = ","))
```

**QC implications:**

- EUR currently distinguishes `ok` intervals from intervals with large altitude changes. These should be retained for profile diagnostics and separated in sensitivity checks.
- CHN intervals currently enter the common layer as `ok`; any CHN-specific stability/QC flags should be added when the QAR level-detection thresholds are finalised.
- The next analysis step should compare all intervals, clean intervals, and duration-filtered intervals before fixing reporting thresholds.

# 4. Lookup Table Framework

## 4.1 Proposed Structure

The refined lookup table provides fuel burn estimates stratified by operational context:

**Dimensions:**
1. **Aircraft group:** Medium (A320/B737 family), Heavy (widebodies)
2. **Distance band:** <500, 500-700, 700-1000, >1000 NM
3. **Phase:** TXO, CLIMB, ENR, DESC, TXI
4. **Altitude band:** <FL100, FL100-FL180, FL180-FL240, >FL240
5. **Profile class:** Short-level, Medium-level, Long-level, No-level

**Metrics per cell:**
- Fuel burn [kg]: P20, P50, P80
- Duration [min]: P20, P50, P80
- kg/min: P20, P50, P80
- kg/NM: P20, P50, P80
- Sample size: n

## 4.2 Example: Medium-Type Enroute Level Intervals

```{r}
#| label: example-lookup
#| tbl-cap: "Example profile lookup cells for medium-type enroute level intervals"
#| warning: false

level_segments |>
  filter(HAS_END, DURATION_SEC > 0, !is.na(TYPE), !is.na(altitude_band)) |>
  mutate(
    aircraft_group = case_when(
      str_detect(TYPE, "A319|A320|A321|A20N|A21N|B737|B738|B739|C919") ~ "Medium",
      TRUE ~ "Other"
    )
  ) |>
  filter(aircraft_group == "Medium", LEVEL_CONTEXT_PHASE == "ENROUTE") |>
  group_by(REGION, altitude_band) |>
  summarise(
    n = n(),
    dur_p20 = quantile(duration_min, 0.2),
    dur_p50 = quantile(duration_min, 0.5),
    dur_p80 = quantile(duration_min, 0.8),
    kg_min_p50 = median(kg_min, na.rm = TRUE),
    .groups = "drop"
  ) |>
  arrange(REGION, altitude_band) |>
  kable(digits = 1,
        col.names = c("Region", "Altitude band", "N", "Duration P20 [min]", "P50 [min]", "P80 [min]", "Median kg/min"),
        format.args = list(big.mark = ","))
```

**Interpretation:**
- Duration percentiles describe the level-off profile itself.
- Median `kg/min` is available in both production tables, but should be treated as exploratory until the phase/segment fuel accounting has been validated side-by-side.
- The table illustrates the intended lookup-table grain: region, aircraft class, phase context, altitude band, and percentile summary.

## 4.3 Validation Strategy

**Internal validation:**
1. Compare P50 values to ICNS2025 aggregated rates (35 kg/min baseline)
2. Check consistency across adjacent distance bands
3. Verify altitude band trends (higher FL = lower kg/NM typically)

**External validation:**
- Operator fuel burn data (confidential, aggregated comparison)
- BADA/AEM model cross-checks
- openAP model comparisons (open-source benchmark)

**Sensitivity analysis:**
- QC flag subsets (all intervals vs ok-only)
- Altitude threshold variations (1000 ft vs 1500 ft)
- Duration filters (exclude <5 min segments)

## 4.4 CHN-EUR Comparison Framework

With both processed datasets now available, the comparison approach is:

1. **Profile matching:** Same distance band, aircraft group, altitude
2. **Percentile comparison:** Compare P20/P50/P80 across regions
3. **Operational context:** Account for airspace design differences
4. **Statistical testing:** Significance of regional differences

**Hypothesis:** Medium-type aircraft show similar fuel burn rates when profile-matched (ICNS2025 suggested 35 kg/min convergence). Refined framework should confirm at finer granularity.

# 5. CHN-EUR Analysis Kick-Off

## 5.1 Available CHN Data Structure

**Source:** QAR/FDR-derived summer 2025 production package from CAAC-OSC operations.

The current R2 handover includes:

- `processed/chn/CHN-canonical-milestones-summer2025.parquet`;
- `processed/chn/CHN-level-segments-summer2025.parquet`;
- `processed/chn/CHN-phase-summaries-summer2025.parquet`.

These products have enough common fields to start profile-matched exploratory analysis against the EUR production products.

## 5.2 Harmonization Requirements

**Items to lock down before final modelling:**

1. **Fuel accounting:** Confirm the segment fuel-burn interpretation in both sources and decide whether to compare kg/min, kg/NM, or phase-normalised totals first.
2. **Aircraft type grouping:** Map EUR ICAO types and CHN aircraft-family strings into a common medium/heavy taxonomy.
3. **Distance bands:** Use `TOTAL_FLOWN_NM` from milestone tables to assign flights to common bands.
4. **Level-segment QC:** Decide how to treat EUR `excessive_alt_change` flags and whether analogous CHN stability flags should be added.
5. **Phase context:** Align `CLIMB`, `ENROUTE`, and `DESCENT` labels with paper phase definitions.

## 5.3 Comparison Dimensions

**Matched profile comparison:**

| Dimension | EUR Data | CHN Data | Match Criteria |
|-----------|----------|----------|----------------|
| Aircraft | A320, B738 | A320, B738 | Same ICAO type |
| Distance | 500-700 NM | 500-700 NM | ±50 NM tolerance |
| Altitude | >FL240 | >FL240 | Same band |
| Phase | Cruise level | Cruise level | Same context |

**Expected outputs:**
- Side-by-side percentile tables
- Regional difference quantification
- Operational context interpretation (airspace design, traffic density)

## 5.4 Immediate Analysis Tasks

The next stepping-stone analysis can now be started directly from the processed data:

1. Build a flight-level profile descriptor table: total distance, aircraft group, number of level segments, total level time, level time by phase, and level time by altitude band.
2. Build matched EUR/CHN lookup cells for aircraft group, distance band, phase context, and altitude band.
3. Add percentile summaries for duration, fuel, kg/min, and kg/NM.
4. Run sensitivity checks for EUR all-segment versus `ok`-only subsets.
5. Compare the resulting P50 and P20 values with the ICNS2025 phase-level baseline.

# 6. Discussion: Towards Operational Lookup Tables

## 6.1 Key Advances Over ICNS2025

**Methodological improvements:**

1. **Profile characterization:** Altitude bands and level-segment analysis replace coarse aggregation
   - Enables fuel burn estimation sensitive to operational context
   - Accounts for stepped climbs, cruise-climb, traffic constraints

2. **Data completeness:** Production EUR and CHN level-segment tables now provide complete interval rows for exploratory analysis
   - EUR FIXED pipeline corrects the earlier imbalance problem
   - CHN QAR products add a second regional sample with comparable canonical fields
   - Transparent QC flags enable sensitivity analysis

3. **Statistical framework:** Percentile-based bounds replace single reference
   - P20/P50/P80 quantify operational variability and benefit pools
   - Enables "what-if" analysis for ANS interventions
   - Supports uncertainty quantification

4. **Reproducibility:** R2 cloud storage with manifest tracking
   - All data products versioned and accessible
   - Processing pipeline documented (technical-note-data-preparation.qmd)
   - Code modular and reusable (R/ function modules)

## 6.2 Data Quality Lessons

**Operational data characteristics:**

- **Level-segment semantics remain source-specific**
  - EUR intervals come from the PRU FIXED extraction and AEM fuel model.
  - CHN intervals come from QAR-derived trajectory and fuel-flow processing.
  - The common analysis table should be treated as a harmonised layer, not proof that source detection algorithms are identical.

- **Altitude stability thresholds need regional calibration**
  - EUR retains `excessive_alt_change` flags for sensitivity analysis.
  - Some flagged intervals may represent operationally meaningful cruise-climb or step-climb behaviour rather than bad data.
  - CHN-specific stability flags should be added if the QAR detection process exposes comparable measures.

- **Phase label consistency varies by data source**
  - EUR and CHN both expose broad level context, but the derivation path differs.
  - Harmonization is essential before interpreting regional differences as operational effects.

**Recommendation for processing teams:**
- Keep the R2 manifest as the handover contract.
- Publish per-release row counts, schema notes, and QC summaries.
- Keep source-specific extraction details separate from harmonised analysis fields.

## 6.3 International Benchmarking Implications

**Harmonization requirements:**

1. **Milestone definitions:** Aligned distance/FL crossing logic documented
2. **Level-segment criteria:** Consistent altitude stability thresholds
3. **QC framework:** Common flag taxonomy (ok, orphaned, excessive_alt_change)
4. **Profile dimensions:** Agreed aircraft groups, distance bands, altitude bands

**Multi-regional validation potential:**

- EUR + CHN baseline (current study)
- Extension to other ICAO regions (NAM, ASIA/PAC, MID)
- Contribution to ICAO GANP KPI16 guidance (additional fuel burn)
- Support for regional performance comparison frameworks

**GANP KPI16 enhancement:**

Current KPI16 converts additional time/distance to fuel burn using average rates. Refined lookup tables enable:

- **Context-specific conversion:** Use profile-matched rates instead of global average
- **Benefit pool quantification:** P50-P20 spread estimates achievable savings
- **Sensitivity bounds:** P80-P20 range for uncertainty
- **Regional comparability:** Harmonized methodology across ANSPs

## 6.4 Limitations and Future Work

**Current limitations:**

1. **Fuel metric interpretation still needs validation**
   - Both production level-segment tables include fuel fields, but the first comparison should confirm like-for-like segment fuel accounting.
   - Cross-check EUR AEM-derived fuel and CHN QAR-derived fuel before making causal claims.

2. **CHN data now available, but harmonisation choices remain**
   - Aircraft type grouping, phase context, and QC flags need a stable comparison taxonomy.
   - Fuel flow unit and conversion checks should remain documented as part of the CHN processing provenance.

3. **Sample representativeness**
   - EUR: Summer 2025 PRU sample across selected European airport pairs
   - CHN: Summer 2025 China Eastern sample across selected airport pairs
   - Seasonal variability not assessed
   - Need multi-year validation

4. **Profile class granularity**
   - Current: distance band + altitude + phase context
   - Future: Add weather (wind, temperature), time-of-day, traffic density
   - Machine learning potential for profile clustering

**Next steps:**

**Short-term (Q4 2026):**
- Validate EUR/CHN segment fuel metrics and compute kg/min and kg/NM consistently
- Generate CHN-EUR profile-matched comparison tables
- Validate P50 values against ICNS2025 baseline (35 kg/min)

**Medium-term (2027):**
- Extend EUR data to multi-year sample (seasonal variability)
- Expand aircraft type coverage (additional medium/heavy types)
- Publish lookup table methodology paper
- Contribute to ICAO GANP KPI16 guidance update

**Long-term (2027+):**
- Multi-regional validation (NAM, ASIA/PAC, MID regions)
- Operational implementation: ANSP fuel burn monitoring dashboards
- Machine learning profile clustering
- Integration with emissions inventories (non-CO₂ effects)

## 6.5 Data Handover Standardization

The current R2 manifest now works as the clean synchronisation point for contributors. The exploratory paper relies only on the current production products:

- `processed/eur/EUR-canonical-milestones-summer2025.parquet`;
- `processed/eur/EUR-level-segments-summer2025.parquet`;
- `processed/chn/CHN-canonical-milestones-summer2025.parquet`;
- `processed/chn/CHN-level-segments-summer2025.parquet`;
- `processed/chn/CHN-phase-summaries-summer2025.parquet`.

Future processing updates should preserve this contract: update the registry, upload the objects, publish the manifest, and verify `r2-sync-status.R` before analysis work proceeds.

# 7. Conclusions

This exploratory document establishes a refined fuel-burn estimation framework addressing limitations of the 2025 ICNS baseline:

**Methodological advances:**
- Level-segment analysis with altitude bands for both EUR and CHN production products
- FIXED EUR processing and newly arrived CHN data provide a clean comparative starting point
- Percentile-based lookup tables (P20/P50/P80 bounds)
- Comprehensive QC framework for data quality transparency

**Initial data insights:**
- Both regions have substantial above-FL240 level-segment activity, but lower-altitude intervals are also material.
- Duration distributions differ by region, phase context, and aircraft mix.
- EUR QC flags should be retained for sensitivity analysis.
- CHN data is now ready for first profile-matched exploratory comparisons.

**Framework status:**
- EUR and CHN production data are loaded from the clean R2 manifest.
- Lookup table structure defined; next step is fuel metric validation and matched-cell analysis.
- Reproducible pipeline with R2 cloud storage

**International benchmarking readiness:**
- Harmonized milestone definitions documented
- QC framework supports multi-regional validation
- GANP KPI16 enhancement potential demonstrated
- Path to operational lookup tables established

**The refined framework provides a foundation for profile-aware fuel burn estimation supporting ICAO net-zero 2050 goals through transparent, reproducible, and internationally comparable performance monitoring.**

---

# References

1. ICAO (2022). Long-Term Aspirational Goal, Special Supplement. [https://www.icao.int/environmental-protection/Documents/EnvironmentalReports/2022/ENVReport2022_Special%20Supplement%20on%20LTAG.pdf](https://www.icao.int/environmental-protection/Documents/EnvironmentalReports/2022/ENVReport2022_Special%20Supplement%20on%20LTAG.pdf)

2. ICAO (2023). GANP KPI Overview. [https://www4.icao.int/ganpportal/ASBU/KPI](https://www4.icao.int/ganpportal/ASBU/KPI)

3. Ma, L., Dai, W., & Koelle, R. (2025). Fuel flow analysis for different flight phases based on a data-driven approach. ICNS 2025 Conference Proceedings.

4. Nuic, A. (2009). Base of Aircraft Data (BADA) Product Management Document. EUROCONTROL.

5. Deransy, R. (2022). EUROCONTROL - AEM, Advanced Emission Model, Flight Emissions Assessment. [https://www.eurocontrol.int/sites/default/files/2022-12/ectl_env_impact_assessment_forum_-_aem.pdf](https://www.eurocontrol.int/sites/default/files/2022-12/ectl_env_impact_assessment_forum_-_aem.pdf)

6. Sun, J., Hoekstra, J., & Ellerbroek, J. (2020). OpenAP: An open-source aircraft performance model for air transportation studies and simulations. *Aerospace*, 7(8), 104. doi:10.3390/aerospace7080104

---

**Document Status:** Exploratory analysis baseline (October 2026)
**EUR Data:** Production FIXED pipeline loaded
**CHN Data:** Summer 2025 production sample loaded
**Next Update:** Profile-matched fuel-burn lookup-table analysis
